### 2016-09-15 Code/2_Examine_LifeExpectancy.Rmd

---
title: "Examining differences in life expectancy across sensitivity analyses"
author: "Corinne Riddell"
date: "July 22, 2016"
output: html_document
---

```{r load_libs, echo=F, warning=F, message=F}
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
```

```{r load_workspace}
load(file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/Sep6_BWgap.Rdata")
```

```{r}

dat3 <- 
dat2 %>% group_by(StateYearRaceSex) %>%
  arrange(Age2) %>%  
  summarise(State2 = first(State2), Year2 = first(Year2), Race2 = first(Race2), Sex2 = first(Sex2),
            Pop_across_age = sum(Population, na.rm = "always"), Pop_has_NA = sum(Population), 
            LE_birth1 = first(life_exp1), LE_birth5 = first(life_exp5), LE_birth9 = first(life_exp9), 
            LE_birthb9 = first(life_expb9), LE_birthw9 = first(life_expw9))

dat3$grouping <- interaction(dat3$State2, dat3$Race2, dat3$Sex2)
dat3$Year <- as.numeric(as.character(dat3$Year2))

#life expectancy over time under two sensitivities for imputing the NA death counts
ggplot(subset(dat3, State2 == "Alabama"), aes(y = LE_birth1, x = Year)) + 
  geom_line(aes(col = grouping, lty = "Imputed 1")) +
  geom_line(aes(y = LE_birth9, col = grouping, lty = "Imputed 9")) +
 # geom_line(aes(y = LE_birthb9, col = grouping, lty = "Imputed w=1, b=9")) +  
  ylab("Life Expectancy at Birth") + scale_x_continuous(breaks = seq(1969, 2013, by = 4)) +
  scale_y_continuous(breaks = seq(55, 90, by = 2)) +
  facet_wrap(~ Sex2)

#graph of population size over time 
#important because gives us an idea of how stable/precise the life expectancy calculation are 
ggplot(subset(dat3, State2 == "Alabama"), aes(y = Pop_across_age, x = Year)) + geom_line(aes(col = grouping)) +
  ylab("Population Size") + scale_y_continuous(label = scales::comma)

ggplot(subset(dat3, State2 == "Massachusetts"), aes(y = LE_birth1, x = Year)) + 
  geom_line(aes(col = grouping, lty = "Imputed 1")) +
  geom_line(aes(y = LE_birth9, col = grouping, lty = "Imputed 9")) +
  geom_line(aes(y = LE_birthb9, col = grouping, lty = "Imputed w=1, b=9")) +  
  ylab("Life Expectancy at Birth") + scale_x_continuous(breaks = seq(1969, 2013, by = 4)) +
  scale_y_continuous(breaks = seq(55, 90, by = 2)) +
  facet_wrap(~ Sex2)

ggplot(subset(dat3, State2 == "Massachusetts"), aes(y = Pop_across_age, x = Year)) + geom_line(aes(col = grouping)) +
  ylab("Population Size") + scale_y_continuous(label = scales::comma)

ggplot(subset(dat3, State2 == "New York"), aes(y = LE_birth1, x = Year)) + 
  geom_line(aes(col = grouping, lty = "Imputed 1")) +
  geom_line(aes(y = LE_birth9, col = grouping, lty = "Imputed 9")) +
  geom_line(aes(y = LE_birthb9, col = grouping, lty = "Imputed w=1, b=9")) +  
  ylab("Life Expectancy at Birth") + scale_x_continuous(breaks = seq(1969, 2013, by = 4)) +
  scale_y_continuous(breaks = seq(55, 90, by = 2)) +
  facet_wrap(~ Sex2)

ggplot(subset(dat3, State2 == "New York"), aes(y = Pop_across_age, x = Year)) + geom_line(aes(col = grouping)) +
  ylab("Population Size") + scale_y_continuous(label = scales::comma)

ggplot(subset(dat3, State2 == "Mississippi"), aes(y = LE_birth1, x = Year)) + 
  geom_line(aes(col = grouping, lty = "Imputed 1")) +
  geom_line(aes(y = LE_birth9, col = grouping, lty = "Imputed 9")) +
  geom_line(aes(y = LE_birthb9, col = grouping, lty = "Imputed w=1, b=9")) +  
  ylab("Life Expectancy at Birth") + scale_x_continuous(breaks = seq(1969, 2013, by = 4)) +
  scale_y_continuous(breaks = seq(55, 90, by = 2)) +
  facet_wrap(~ Sex2)

ggplot(subset(dat3, State2 == "Mississippi"), aes(y = Pop_across_age, x = Year)) + geom_line(aes(col = grouping)) +
  ylab("Population Size") + scale_y_continuous(label = scales::comma)
```

```{r}
dat4_Black <- dat3[dat3$Race2 == "Black", ]
dat4_Black <- dat4_Black %>% 
  select(State2, Year, Sex2, Pop_across_age, Pop_has_NA, LE_birth1, LE_birth5, LE_birth9) %>%
  rename(Pop_black = Pop_across_age, Pop_NA_black = Pop_has_NA, LE1_black = LE_birth1, LE5_black = LE_birth5, LE9_black = LE_birth9)

dat4_White <- dat3[dat3$Race2 == "White", ]
dat4_White <- dat4_White %>% 
  select(State2, Year, Sex2, Pop_across_age, Pop_has_NA, LE_birth1, LE_birth5, LE_birth9) %>%
    rename(Pop_white = Pop_across_age, Pop_NA_white = Pop_has_NA, LE1_white = LE_birth1, LE5_white = LE_birth5, LE9_white = LE_birth9)

dat4 <- merge(dat4_Black, dat4_White, by = c("State2", "Year", "Sex2"))
dim(dat4)
View(dat4)

dat4 <- dat4 %>% mutate(gap_w1b1 = LE1_white - LE1_black, 
                gap_w5b5 = LE5_white - LE5_black,
                gap_w9b9 = LE9_white - LE9_black,
                gap_w1b9 = LE1_white - LE9_black,
                gap_w9b1 = LE9_white - LE1_black)

summary(dat4$Pop_white)
summary(dat4$Pop_black)

#Life expectancy gap, under the different sensitivity analyses
ggplot(subset(dat4, State2 == "Alabama"), aes(x = Year, y = gap_w1b1)) + 
  geom_line(aes(y = gap_w1b9, col = "Imputed w=1, b=9")) +   
  geom_line(aes(col = "Imputed 1")) + 
  geom_line(aes(y = gap_w5b5, col = "Imputed 5")) +
  geom_line(aes(y = gap_w9b9, col = "Imputed 9")) +
  geom_line(aes(y = gap_w9b1, col = "Imputed w=9, b=1")) +
  facet_wrap( ~ Sex2) +
  ylab("Life Expectancy Gap") + 
  scale_x_continuous(breaks = seq(1969, 2013, by = 4)) +
  scale_y_continuous(breaks = seq(-1, 11, by = 1))
```