### 2016-09-22 Code/error checking the LE decomp fxs.Rmd
improved LE functions and split files for testing

---
title: "error checking the LE decomp functions"
author: "Corinne Riddell"
date: "September 22, 2016"
output: html_document
---

In this file we apply the functions created to calculate and decompose life expectancy by way of example, and as a way to cross-check against results from published findings (to ensure my functions work properly).

Our functions:
```{r source_LE_fxs}
library(dplyr)
library(ggplot2)
source("/Users/corinneriddell/Documents/repos/BlackWhiteGap/Code/life_expectancy_functions.R")
```

**First example**

We use data from a paper published by Auger et al and compare our findings to those found by the authors.

Here is the data need to calculate life expectancy, as provided by the authors in a supplementary appendix (xls file) to their paper:
```{r create_dataset}
canada.men <- data.frame("ages" = c(0, 1, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90),
                         "obs.deaths" = c(4054, 625, 445, 622, 2568, 3770, 3434, 3681, 5155, 8449, 13421, 19371, 
                                          25448, 31465, 37441, 48796, 64763, 72512, 60109, 43812),
                         "obs.popn" = c(719953, 2835604, 3654546, 4071400, 4464549, 4564679, 4296613, 4237402,
                                        4532151, 5085092, 5163532, 4597916, 3937984, 3120981, 2323629, 1862971,
                                        1499907, 997492, 492445, 196109))

quebec.men <- data.frame("ages" = c(0, 1, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90), 
                         "obs.deaths" = c(1057, 164, 116, 176, 659, 956, 1056, 1110, 1498, 2470, 4116, 6054, 8426,
                                          11175, 13558, 16567, 21349, 22396, 16851,10878),
                         "obs.popn" = c(215200, 789710, 1008266, 1179423, 1249360, 1261780, 1349857, 1293521,
                                        1313834, 1551684, 1622835, 1485249, 1288014, 1077253, 796557, 609341, 
                                        476211, 300506, 134353, 49678))
```

I then add a column that denotes how many years are in each age bracket:
```{r add_n_x}
canada.men$n_x <- 5
canada.men$n_x[canada.men$ages == 0] <- 1
canada.men$n_x[canada.men$ages == 1] <- 4

quebec.men$n_x <- 5
quebec.men$n_x[quebec.men$ages == 0] <- 1
quebec.men$n_x[quebec.men$ages == 1] <- 4

head(canada.men)
head(quebec.men)
```

We also need data on the proportion of deaths within each age bracket due to each cause of death.
```{r data_cod_proportions}
cod.props <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteGap/working_teaching_examples/auger_cod_proportions.csv")
names(cod.props)[3] <- "prop.QC"
names(cod.props)[4] <- "prop.CAD"
head(cod.props)
```

Calculate and display the life tables for the Canada and Quebec data:
```{r calculate_life_expectancy}
life.table.cm <- life.table(data = canada.men, age.groups = "ages", num.ages.in.group = "n_x", 
                            death.counts = "obs.deaths", population.counts = "obs.popn")

round(life.table.cm, 2)

life.table.qm <- life.table(data = quebec.men, age.groups = "ages", num.ages.in.group = "n_x", 
                            death.counts = "obs.deaths", population.counts = "obs.popn")

round(life.table.qm, 2)
```

Life expectancy at birth in Canadian men is `r round(life.table.cm$e_x[1], 2)` vs. `r round(life.table.qm$e_x[1], 2)` in Quebec men, for a life expectancy gap of `r life.table.cm$e_x[1] - life.table.qm$e_x[1]` years.

Decompose the life expectancy gap by age:

```{r calc_age_decomp}
age_decomp <- le_age_decomp(life.table1 = life.table.qm, name.lt1 = "Quebec", 
                            life.table.cm, name.lt2 = "Canada", 
                            age.groups = "ages")
age_decomp
sum(age_decomp$C_x) 
```

Check that the function still works when the life tables are provided in the reverse order:

```{r calc_age_decomp_reverse}
age_decomp2 <- le_age_decomp(life.table1 = life.table.cm, name.lt1 = "Canada", 
                             life.table2 = life.table.qm, name.lt2 = "Quebec", 
                             age.groups = "ages")
age_decomp2
sum(age_decomp2$C_x)
```

Compare the output of the above two functions -- the contribution of age to the LE gap are slightly different based on the order of operation:

```{r compare decomp_output}
round(cbind(age_decomp$C_x, age_decomp2$C_x),4)
```

```{r calc_cod_decomp}
canada.cod.decomp <- cause_of_death_decomp(life.table1 = life.table.qm, life.table2 = life.table.cm, decomp.table = age_decomp,
                      cod.table = cod.props, age.colname.cod.table = "Age", COD.colname.cod.table = "cod", 
                      prop1.colname.cod.table = "prop.QC", prop2.colname.cod.table = "prop.CAD")
```

Collapse by cause of death to calculate the overall contribution of each cause of death to the gap in life expectnancy:
```{r }
cod.contribution <- canada.cod.decomp %>% group_by(Cause.of.death) %>% summarise(cod.contribution = sum(C_xi))

#order by contribution amount for graphing purposes
cod.contribution$Cause.of.death <- factor(cod.contribution$Cause.of.death, 
                                          levels = cod.contribution$Cause.of.death[order(cod.contribution$cod.contribution,
                                                                                         decreasing = T)])

ggplot(cod.contribution, aes(x = Cause.of.death, y = cod.contribution)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_abline(intercept = 0, slope = 0)
```

**A second example**

In this example we use the data that Sam entered into Stata to illustrate the life expectancy concept to me. I'll replicate the life table here to ensure that I can arrive at the same results as he does in Stata.

```{r lifetable_cali_complete}
cali.complete <- data.frame("ages" = c(0:15), "nx" = rep(1,16),
                         "obs.deaths" = c(1635, 64, 41, 22, 41, 55, 42, 58, 44, 52, 48, 60, 52, 82, 129, 23300), #last number is fictitious
                         "obs.popn" = c(123342, 111520, 109200, 108749, 105698, 110548, 106857, 112184, 116423,
                                        132952, 134266, 128938, 125502, 128212, 132775, 143600))

head(cali.complete)

life.table(data = cali.complete, age.groups = "ages", num.ages.in.group = "nx", 
           death.counts = "obs.deaths", population.counts = "obs.popn")

cali.complete$avelived <- 0.5
cali.complete$avelived[cali.complete$ages==0] <- 0.1

life.table(data = cali.complete, age.groups = "ages", num.ages.in.group = "nx", ave.prop.lived = "avelived",
           death.counts = "obs.deaths", population.counts = "obs.popn")
```

```{r lifetable_cali_abridged, warning=F, message=F}
cali.complete$group <- cut(cali.complete$ages, breaks=c(0, 1, 5, 10, 15, 20), 
                           labels = c("<1 years old", "1-4 years old", "5-9 years old", 
                                      "10-14 years old", "15 or older"), include.lowest = T, right = F)

table(cali.complete$group, cali.complete$ages)

library(dplyr)
cali.abridged <- cali.complete %>% group_by(group) %>% 
                                   summarise(nx = sum(nx), ages = first(group), popn = sum(obs.popn), deaths = sum(obs.deaths),
                                             avelived = first(avelived))

cali.abridged

data.frame(life.table(data = cali.abridged, age.groups = "ages", num.ages.in.group = "nx", 
           death.counts = "deaths", population.counts = "popn"))

data.frame(life.table(data = cali.abridged, age.groups = "ages", num.ages.in.group = "nx", 
           death.counts = "deaths", population.counts = "popn", ave.prop.lived = "avelived"))
```