### 2016-09-28 Code/Test life expectancy fxs on Alabama.Rmd
more updates

---
title: "Test functions on Alabama data"
author: "Corinne Riddell"
date: "September 22, 2016"
output: html_document
---

```{r initialize}
library(dplyr)
library(ggplot2)
library(plotly)
load("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/alabama_only.Rdata")
#example data
```

```{r}
black.males.alabama <- alabama.by.strata[alabama.by.strata$Race2 == "Black" &
                                           alabama.by.strata$Sex2 == "Male", ]

males.alabama <- alabama.by.strata[ alabama.by.strata$Sex2 == "Male", ]

ggplot(black.males.alabama, aes(x = Year , y = Age3, z = log(mortality_rate, 10))) + 
 # stat_contour(geom = "polygon", aes(fill = ..level..), bins = 19) 
#         geom_tile(aes(fill = mortality_rate)) 
         stat_contour(bins = 10, aes(colour = ..level..) )

plot_ly(data = black.males.alabama, x = Year, y = Age3, z = log(mortality_rate,10), type = "heatmap")

alabama.by.strata$id <- as.integer(alabama.by.strata$RaceSex)

test <- plot_ly(data = alabama.by.strata, x = Year, y = Age3, z = log(mortality_rate,10),
        group = RaceSex, type = "heatmap", xaxis = paste0("x", id))

subplot(test)

ggplot(data = alabama.by.strata, aes(x=Year2, y = Age3)) + geom_tile(aes(fill = log(mortality_rate,10))) +
  facet_grid(Race2 ~ Sex2)

first.strata.cod2 <- alabama.by.strata.cod[alabama.by.strata.cod$RaceSex == "White.Male", ]
first.strata.cod2$Year3 <- as.numeric(as.character(first.strata.cod2$Year2))

plot_ly(data = subset(first.strata.cod2, COD2 == "Injuries"), x = Year3, y = Age3, z = cod_prop_death5, type = "heatmap")

plot_ly(data = subset(first.strata.cod2, COD2 == "Communicable"), x = Year3, y = Age3, z = cod_prop_death5, type = "heatmap")

plot_ly(data = subset(first.strata.cod2, COD2 == "Cancers"), x = Year3, y = Age3, z = cod_prop_death5, type = "heatmap")

plot_ly(data = subset(first.strata.cod2, COD2 == "Non-communicable"), x = Year3, y = Age3, z =cod_prop_death5, type = "heatmap")

#this is wrong - need to create the dataset...
alabama.by.strata.cod2$id <- as.integer(alabama.by.strata.cod2$RaceSex)
subplot(plot_ly(data = subset(alabama.by.strata.cod2, COD2 == "Communicable"), x = Year3, y = Age3, z = cod_prop_death5, type = "heatmap", group = RaceSex, xaxis = paste0("x", id)))

first.strata.cod$id <- as.integer(first.strata.cod$COD2)
plot_ly(data = first.strata.cod, x = Year, y = Age3, z = log(cod_prop_death5, 10), type = "heatmap", group = COD2,
        xaxis = paste0("x", id))
```

```{r alabama_example}
alabama.by.strata <- 
  dat.clean.alabama %>% group_by(Year2, RaceSex, Age2) %>%
  arrange(RaceSex, Year2, Age2) %>%
  summarise(total_deaths5 = sum(Count_md5),
            Population = first(Population), 
            Race2 = first(Race2), 
            Sex2 = first(Sex2), 
            Year = first(Year),
            Age = first(Age), Age3 = first(Age3),
            StateYearRaceSex = first(StateYearRaceSex),
            nx = first(nx)) %>%
  mutate(mortality_rate = total_deaths5/Population)

first.strata <- alabama.by.strata[alabama.by.strata$StateYearRaceSex == "Alabama.1969.White.Male", ]

second.strata <- alabama.by.strata[alabama.by.strata$StateYearRaceSex == "Alabama.1969.Black.Male", ]

alabama.by.strata.cod <- dat.clean.alabama %>% group_by(Year2, RaceSex, Age2) %>%
                                               arrange(RaceSex, Year2, Age2) %>%
                                               mutate(total_deaths5 = sum(Count_md5),
                                                      cod_prop_death5 = Count_md5/total_deaths5)

first.strata.cod <- alabama.by.strata.cod[alabama.by.strata.cod$StateYearRaceSex == "Alabama.1969.White.Male", ]
second.strata.cod <- alabama.by.strata.cod[alabama.by.strata.cod$StateYearRaceSex == "Alabama.1969.Black.Male", ]

cod.tab.alabama.male.1969 <- first.strata.cod %>% ungroup() %>%
                                                  select(Age2, COD2, cod_prop_death5) %>%
                                                  rename(prop2 = cod_prop_death5)

cod.tab.alabama.male.1969$prop1 <- second.strata.cod$cod_prop_death5
cod.tab.alabama.male.1969 <- cod.tab.alabama.male.1969 %>% arrange(COD2, Age2)

ggplot(cod.tab.alabama.male.1969, aes(x=COD2, y=prop1)) + 
  geom_point(aes(col="White")) + geom_point(aes(y=prop2, col="Black")) +
  scale_y_continuous(name = "Proportion dying from specified cause") +
  facet_wrap(~Age2) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

system.time(life.table.whites <- data.frame(life.table(data = first.strata, age.groups = "Age2", num.ages.in.group = "nx", death.counts = "total_deaths5", population.counts = "Population")))

life.table.blacks <- data.frame(life.table(data = second.strata, age.groups = "Age2", num.ages.in.group = "nx", death.counts = "total_deaths5", population.counts = "Population"))

ggplot(life.table.blacks, aes(x = Age2, y = q_x)) + geom_point() + geom_point(aes(y = R_x), col="red")

life.table.whites[, c("Age2", "total_deaths5", "Population", "R_x", "q_x", "l_x", "d_x", "L_x", "T_x", "e_x")]

system.time(
for(stratum.id in levels(droplevels(alabama.by.strata$StateYearRaceSex))) {
  assign(paste0("lt.", as.character(stratum.id)), 
         life.table(data = subset(alabama.by.strata, StateYearRaceSex == stratum.id), 
                                  age.groups = "Age2", num.ages.in.group = "nx", 
                                  death.counts = "total_deaths5", population.counts = "Population"))
}
)

alabama_age_decomp <- le_age_decomp(life.table.blacks, life.table.whites, "Age2")
alabama_age_decomp

ggplot(alabama_age_decomp, aes(y = C_x, x = Ages)) + geom_point() + scale_y_continuous("Contribution of Age to Gap")
sum(alabama_age_decomp$C_x)
life.table.whites$e_x[1] - life.table.blacks$e_x[1]

ggplot(alabama_age_decomp, aes(y = Life_Expectancy_1, x = Ages)) + geom_point(aes(col="Black")) + scale_y_continuous("Life Expectancy (Yrs)") +
  geom_point(aes(y = Life_Expectancy_2, col="White"))

alabama_cod_decomp <- cause_of_death_decomp(life.table1 = life.table.blacks, life.table2 = life.table.whites, decomp.table = alabama_age_decomp,
                      cod.table = cod.tab.alabama.male.1969, "Age2", "COD2", "prop1", "prop2")
```