### 2016-11-22 Code/Explore_unaggregate_data.Rmd
Create AL cut of unaggregated dataset

---
title: "Untitled"
author: "Corinne Riddell"
date: "November 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(plotly)
source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")
```

```{r load datasets}
#loading in the data from NCHS, which supresses death counts between 1 and 9.
dat.male <- read.table("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/males_export.txt", header = T, sep="\t", na.strings = " ")
names(dat.male) <- c("State", "Age", "Sex", "Race", "COD", "Year", "Crude_Rate", "Count", "Population")
prop.table(table(is.na(dat.male$Count)))
head(dat.male)
table(dat.male$Sex)
str(dat.male) #note all varinables are numeric or integer

dat.female <- read.table("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/female_export.txt", header = T, sep="\t", na.strings = " ")
names(dat.female) <- c("State", "Age", "Sex", "Race", "COD", "Year", "Crude_Rate", "Count", "Population")
table(dat.female$Sex)
dat.female$Sex <- 1
str(dat.male) #note all varinables are numeric or integer

prop.table(table(is.na(dat.female$Count)))
head(dat.female)

dat.both <- rbind(dat.male, dat.female)

rm(dat.male, dat.female)
```

```{r load_COD_dictionary}
cod.dict <- read.csv("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/COD_dictionary.csv")

cod.dict$COD2 <- reorder(cod.dict$COD, cod.dict$COD_num)
levels(cod.dict$COD2)

cod.dict$cod_cat <- "Cancers"

cod.dict$cod_cat[cod.dict$COD2 == "Diseases of Heart" | 
                 cod.dict$COD2 == "Hypertension without Heart Disease" |
                 cod.dict$COD2 == "Cerebrovascular Diseases" |  
                 cod.dict$COD2 == "Atherosclerosis" |  
                 cod.dict$COD2 == "Aortic Aneurysm and Dissection" | 
                 cod.dict$COD2 == "Other Diseases of Arteries, Arterioles, Capillaries"] <- "Cardiovascular"

cod.dict$cod_cat[cod.dict$COD2 == "Tuberculosis" | 
                 cod.dict$COD2 == "Syphilis" |
                 cod.dict$COD2 == "Human Immunodeficiency Virus (HIV) (1987+)" |  
                 cod.dict$COD2 == "Septicemia" |  
                 cod.dict$COD2 == "Other Infectious and Parasitic Diseases" | 
                 cod.dict$COD2 == "Pneumonia and Influenza"] <- "Communicable"

cod.dict$cod_cat[cod.dict$COD2 == "Diabetes Mellitus" | 
                 cod.dict$COD2 == "Alzheimers (ICD-9 and 10 only)" |
                 cod.dict$COD2 == "Chronic Obstructive Pulmonary Disease and Allied Cond" |  
                 cod.dict$COD2 == "Stomach and Duodenal Ulcers" |  
                 cod.dict$COD2 == "Chronic Liver Disease and Cirrhosis" | 
                 cod.dict$COD2 == "Nephritis, Nephrotic Syndrome and Nephrosis"] <- "Non-communicable"

cod.dict$cod_cat[cod.dict$COD2 == "Accidents and Adverse Effects" | 
                 cod.dict$COD2 == "Suicide and Self-Inflicted Injury" |
                 cod.dict$COD2 == "Homicide and Legal Intervention"] <- "Injuries"

cod.dict$cod_cat[cod.dict$COD2 == "Complications of Pregnancy, Childbirth, Puerperium" | 
                 cod.dict$COD2 == "Congenital Anomalies" |
                 cod.dict$COD2 == "Certain Conditions Originating in Perinatal Period" |  
                 cod.dict$COD2 == "Symptoms, Signs and Ill-Defined Conditions" |  
                 cod.dict$COD2 == "Other Cause of Death"] <- "All other causes"

cod.dict <- cod.dict[ ,-2]
```

```{r recode_variables}
dat.both <- dat.both %>%
  mutate(State2 = factor(State, levels = c(0:50), labels = c("Alabama", "Alaska", "Arizona", "Arkansas","California",
                                                             "Colorado", "Connecticut", "Delaware", "Washington DC",
                                                             "Florida", "Georgia", "Hawaii","Idaho", "Illinois", 
                                                             "Indiana","Iowa", "Kansas", "Kentucky", "Louisiana", 
                                                             "Maine", "Maryland", "Massachusetts", "Michigan", 
                                                             "Minnesota", "Mississippi","Missouri", "Montana", 
                                                             "Nebraska", "Nevada", "New Hampshire", "New Jersey",
                                                             "New Mexico", "New York","North Carolina","North Dakota",
                                                             "Ohio", "Oklahoma","Oregon", "Pennsylvania", 
                                                             "Rhode Island", "South Carolina", "South Dakota", 
                                                             "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
                                                             "Washington", "West Virginia", "Wisconsin", "Wyoming")), 
         Age2 = factor(Age, levels = c(0:19), labels = c("<1 year", "1-4 years", "5-9 years", "10-14 years", 
                                                         "15-19 years", "20-24 years", "25-29 years", "30-34 years",
                                                         "35-39 years", "40-44 years", "45-49 years", "50-54 years", 
                                                         "55-59 years", "60-64 years", "65-69 years", "70-74 years", 
                                                         "75-79 years", "80-84 years","85+ years", "Unknown"), ordered = T),
         Age3 = 0*(Age2 == "<1 year") + 1*(Age2 == "1-4 years") + 5*(Age2 == "5-9 years") +
                10*(Age2 == "10-14 years") + 15*(Age2 == "15-19 years") + 20*(Age2 == "20-24 years") + 
                25*(Age2 == "25-29 years") + 30*(Age2 == "30-34 years") + 35*(Age2 == "35-39 years") +
                40*(Age2 == "40-44 years") + 45*(Age2 == "45-49 years") + 50*(Age2 == "50-54 years") + 
                55*(Age2 == "55-59 years") + 60*(Age2 == "60-64 years") + 65*(Age2 == "65-69 years") +   
                70*(Age2 == "70-74 years") + 75*(Age2 == "75-79 years") + 80*(Age2 == "80-84 years") + 
                85*(Age2 == "85+ years"),
         nx = 1*(Age2 == "<1 year") + 4*(Age2 == "1-4 years") + 5*(Age2 != "<1 year" & Age2 != "1-4 years"),
         Sex2 = factor(Sex, levels = c(0, 1), labels = c("Male", "Female")),
         Race2 = factor(Race, levels = c(0, 1), labels = c("White", "Black")),
         COD2 = factor(COD, levels = c(0:92), labels = levels(cod.dict$COD2)), 
         Year2 = factor(Year, levels = c(0:44), labels = as.character(1969:2013)),
         Year3 = as.numeric(as.character(Year2)),
         RaceSex = interaction(Race2, Sex2, sep = " "),
         StateYearSex = interaction(State2, Year2, Sex2),
         StateYearRaceSex = interaction(State2, Year2, Race2, Sex2))

head(dat.both)

dat.both <- dat.both[dat.both$Age2 != "Unknown", ]

dat.both <- dat.both %>% mutate(stabbrs = factor(State, levels = c(0:50), 
                                      labels = c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC",
                                                 "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", 
                                                 "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", 
                                                 "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", 
                                                 "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", 
                                                 "VT", "VA", "WA", "WV", "WI", "WY")))

dat.both$agegrp <- dat.both$Age 
dat.both$agegrp[dat.both$Age == 0]  <- 1

dim(dat.both)
dat.both <- merge(dat.both, cod.dict, by = "COD2") #to add on COD_num and cod_cat
dim(dat.both)
```

```{r for_merge}
years0610 <- dat.both[dat.both$Year3 >= 2006 & dat.both$Year3 <= 2010 & dat.both$agegrp > 1, ] 
#added the age restriction because don't want to deal with the differences in age definition right now

not.suppressed <- foreign::read.dta("/Users/corinneriddell/Dropbox/VitalStats/notsuppressed_0610.dta")

not.suppressed <- not.suppressed %>% 
  filter(agegrp > 1)

years0610 <- years0610 %>% rename(COD_full = COD2)

together <- merge(years0610, not.suppressed, by = c("stabbrs", "Year3", "Sex", "Race", "agegrp", "COD_full"))
```

```{r}
together <- together %>% mutate(diff_count = Count - death, 
                                diff_pop = Population - popestimate,
                                rel_diff_pop = (Population - popestimate)/Population)

prop.table(table(together$death[is.na(together$Count) == T]))

ggplot(together, aes(x = diff_count)) + geom_density()
summary(together$diff_count)

ggplot(together, aes(x = diff_pop)) + geom_density()
summary(together$diff_pop)

ggplot(together, aes(x = rel_diff_pop)) + geom_density()
summary(together$rel_diff_pop)

View(together[together$rel_diff_pop > 0.6, ])
```
From above we see that the death counts match exactly and when the original count is NA the non-suppressed count is between 1 and 9.
For some reason there is some discrepancy in the population denominators. These do not match exactly and I don't know why.

Since we are interested in the effect of imputing the NAs and not the effect of the different denominators, we need to use the same denominators in the analysis.

```{r pull_alabama}
alabama.males <- together %>% filter(State2 == "Alabama", Sex2 == "Male") 
save(alabama.males, file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/alabama_males_5plus_granular.Rdata")
```