### 2016-11-24 Code/append_smoothed_results.Rmd

---
title: "Append_smoothed_results"
author: "Corinne Riddell"
date: "November 5, 2016"
output: html_document
---

```{r load_libraries}
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(plotly)
source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")
```

```{r load_datasets}
load(file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/Sep6_BWgap.Rdata")
ktm_smoothed <- read.csv("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/smoothed_results.csv")
ktm_rest <- read.csv("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/smoothed_results_extra_states.csv")
ktm_virginia_males <- read.csv("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/results_males_Virginia_Nov21.csv")

dat.clean <- dat.clean[, -c(31:40)]

head(ktm_smoothed)
head(ktm_rest)
head(ktm_virginia_males)

table(ktm_rest$state == "Virginia")
table(ktm_rest$state)
table(ktm_smoothed$state) #check - i think Mass is in both rest and smoothed so need to remove duplicates.

dim(ktm_rest)
ktm_rest <- ktm_rest %>% filter(state != "Massachusetts") #delete duplicate
dim(ktm_rest)

length(ktm_rest$sex[ktm_rest$state ==  "Virginia" & ktm_rest$sex == "Male"])
dim(ktm_rest)
ktm_rest <- ktm_rest %>% filter(!(state == "Virginia" & sex == "Male"))
dim(ktm_rest)

ktm_virginia_males <- ktm_virginia_males[ , -1]

ktm_all <- rbind(ktm_smoothed, ktm_rest, ktm_virginia_males)
```

```{r clean_smoothed_data}
ktm_clean <- ktm_all %>% 
  rename(Race2 = race, Sex2 = sex, State2 = state, COD2 = cod) %>%
  mutate(Age2 = factor(age_bin, levels = c(1:19), 
                       labels = c("<1 year", "1-4 years", "5-9 years", "10-14 years", 
                                  "15-19 years", "20-24 years", "25-29 years", "30-34 years",
                                  "35-39 years", "40-44 years", "45-49 years", "50-54 years", 
                                  "55-59 years", "60-64 years", "65-69 years", "70-74 years", 
                                  "75-79 years", "80-84 years","85+ years"), ordered = T),
         Year2 = factor(year, levels = c(1:45), labels = as.character(1969:2013))
         )
```

```{r merge_results}
dim(dat.clean)
dat.clean <- merge(dat.clean, ktm_clean, 
                   by = c("State2", "Year2", "Race2", "Sex2", "Age2", "COD2"), 
                   all.x = T)
dim(dat.clean)
dat.clean <- dat.clean %>% mutate(Appended_smoothed = ifelse(is.na(X), F, T))

table(dat.clean$Appended_smoothed)
prop.table(table(dat.clean$Appended_smoothed))
```

Ensure that the merge was correct by comparing the duplicate variables (i.e., those variables in both datasets that had different names)

```{r check_merge}
dat.clean[1:10, c("smoothed_deaths", "deaths", "Count", "population", "Population", "State2", "Age2", "Sex2", "Race2", "COD2")]

dat.clean <- dat.clean %>% mutate(diff.pop = Population - population, diff.count = Count - deaths)
summary(dat.clean$diff.pop)
summary(dat.clean$diff.count)
```

Compare the smoothed_results to the original counts to see if they seem reasonable:

```{r compared_original_vs_smoothed}
ggplot(dat.clean, aes(x = Count, y = smoothed_deaths)) + geom_point(alpha = 0.1)
```

Can't really tell anything from this -- make another graphic:

```{r compared_original_vs_smoothed2}
dat.clean <- dat.clean %>% mutate(diff.orig.smooth = Count - smoothed_deaths)
summary(dat.clean$diff.orig.smooth)
ggplot(dat.clean, aes(x = diff.orig.smooth)) + geom_density() + geom_rug(alpha = 0.1)

ggplot(subset(dat.clean, State2 == "Georgia"), aes(x = diff.orig.smooth)) + geom_histogram() + geom_rug(alpha = 0.1) + facet_grid(Sex2~Race2)
ggplot(subset(dat.clean, State2 == "Georgia" & Sex2 == "Male" & Race2 == "Black"), aes(x = diff.orig.smooth)) + geom_histogram() + geom_rug(alpha = 0.1) + facet_wrap(~Age3)

ggplot(subset(dat.clean, State2 == "California"), aes(x = diff.orig.smooth)) + geom_histogram() + geom_rug(alpha = 0.1) + facet_grid(Sex2~Race2)
ggplot(subset(dat.clean, State2 == "California" & Sex2 == "Male" & Race2 == "Black"), aes(x = diff.orig.smooth)) + geom_histogram() + geom_rug(alpha = 0.1) + facet_wrap(~Age3)

ggplot(subset(dat.clean, State2 == "Utah"), aes(x = diff.orig.smooth)) + geom_density() + geom_rug(alpha = 0.1) + facet_grid(Sex2~Race2)
ggplot(subset(dat.clean, State2 == "Utah" & Sex2 == "Male" & Race2 == "Black"), aes(x = diff.orig.smooth)) + geom_histogram() + geom_rug(alpha = 0.1) + facet_wrap(~Age3)
```

```{r addvar_cod_proportions}
dat.clean <- dat.clean %>% group_by(State2, RaceSex, Year3, Age3) %>%
                           arrange(State2, RaceSex, Year3, Age3) %>%
                           mutate(total_smoothed_deaths = sum(smoothed_deaths),
                                  cod_prop_smoothed_deaths = smoothed_deaths/total_smoothed_deaths
                                  )
summary(dat.clean$cod_prop_smoothed_deaths[dat.clean$Appended_smoothed == F])

length(dat.clean$total_smoothed_deaths[dat.clean$Appended_smoothed == T])
length(dat.clean$total_smoothed_deaths[dat.clean$total_smoothed_deaths == 0 & dat.clean$Appended_smoothed == T])
#the smoothed results are NEVER exactly equal to 0.

#there are quite a few < 1:
length(dat.clean$total_smoothed_deaths[dat.clean$total_smoothed_deaths < 0.1 & dat.clean$Appended_smoothed == T])
length(dat.clean$total_smoothed_deaths[dat.clean$total_smoothed_deaths < 0.5 & dat.clean$Appended_smoothed == T])
length(dat.clean$total_smoothed_deaths[dat.clean$total_smoothed_deaths < 1 & dat.clean$Appended_smoothed == T])

View(dat.clean[dat.clean$total_smoothed_deaths < 1 & dat.clean$Appended_smoothed == T, ])
```

```{r aggregate_over_COD}
dat.smoothed.aggregatedCOD <- dat.clean %>% 
  group_by(State2, Year2, RaceSex, Age2) %>%
  arrange(State2, RaceSex, Year2, Age2) %>%
  filter(Appended_smoothed == T) %>%
  summarise(Population = first(Population),
            Race2 = first(Race2), 
            Sex2 = first(Sex2), 
            StateYearRaceSex = first(StateYearRaceSex),
            nx = first(nx),
            Age3 = first(Age3),
            Year3 = first(Year3),
            total_smoothed_deaths = first(total_smoothed_deaths),
            cod_prop_smoothed_deaths = first(cod_prop_smoothed_deaths))

dim(dat.smoothed.aggregatedCOD)

non.zero.stratums.smoothed <-  unique(dat.smoothed.aggregatedCOD$StateYearRaceSex)
```

In those strata that KTM ran the model, create a life table and store it in a list.

```{r calculate_life_expectancy}
#15 mins

list.life.tables.smoothed <- vector(mode = "list", length = length(non.zero.stratums.smoothed))
list.index <- 1

system.time(
for(stratum.id in non.zero.stratums.smoothed) {

  list.life.tables.smoothed[[list.index]] <- life.table(data = subset(dat.smoothed.aggregatedCOD, StateYearRaceSex == stratum.id), 
                                               num.ages.in.group = "nx", 
                                               death.counts = "total_smoothed_deaths", 
                                               population.counts = "Population")
  list.index <- list.index + 1
  print(list.index)
  }
)
```

Add the life expectancy at birth to the dat2 data file:

```{r add_le_to_dataset}
le.smoothed <- rep(NA, length(non.zero.stratums.smoothed))

for (i in 1:length(non.zero.stratums.smoothed)) {
  le.smoothed[i] <- list.life.tables.smoothed[[i]]$e_x[1]
}

les.smoothed <- data.frame(life.expectancy.birth.smoothed = le.smoothed, StateYearRaceSex = non.zero.stratums.smoothed)

dim(dat2)
dat3 <- merge(dat2, les.smoothed, by = "StateYearRaceSex", all.x =T)
dim(dat3)

dat2 <- dat3
rm(les.smoothed, dat3)
```

```{r create_dataset_BlackWhiteSmooth}
#this dataset contains the black-white data within the same row (wide format) for each year-state-sex strata

BlackSmooth <- dat2 %>% filter(Race2 == "Black" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, life.expectancy.birth.smoothed) %>%
                  rename(le.smoothed.black = life.expectancy.birth.smoothed)

WhiteSmooth <- dat2 %>% filter(Race2 == "White" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, life.expectancy.birth.smoothed) %>%
                  rename(le.smoothed.white = life.expectancy.birth.smoothed)

BlackWhiteSmooth <- merge(BlackSmooth, WhiteSmooth, by = c("State2", "Year3", "Sex2"))

BlackWhiteSmooth <- BlackWhiteSmooth %>% mutate(WBgap.s = le.smoothed.white - le.smoothed.black)

dim(BlackWhite)
BlackWhite3 <- merge(BlackWhite, BlackWhiteSmooth, by = c("State2", "Year3", "Sex2"))
dim(BlackWhite3)

BlackWhite <- BlackWhite3
rm(BlackWhite3, BlackSmooth, WhiteSmooth)
```

```{r calculate_age_decomp_smoothed}
#20 mins

list.age.decomp.tables.smoothed <- vector(mode = "list", length = dim(paired.ids)[1])

for(i in 1:dim(paired.ids)[1]) {
  white.index <- which(non.zero.stratums.smoothed == paired.ids[i, "StateYearRaceSexWhite"])
  black.index <- which(non.zero.stratums.smoothed == paired.ids[i, "StateYearRaceSexBlack"])

  if(length(white.index) != 0 & length(black.index) != 0) {
    list.age.decomp.tables.smoothed[[i]] <- le_age_decomp(life.table1 = as.data.frame(list.life.tables.smoothed[[black.index]]),
                                                          name.lt1 = "Black", 
                                                          life.table2 = as.data.frame(list.life.tables.smoothed[[white.index]]),
                                                          name.lt2 = "White", age.groups = "Age2")
  } else {
    list.age.decomp.tables.smoothed[[i]] <-"Couldn't compare life tables"
  }
  print(i)
}
```

```{r calculate_cod_decomp}
#takes 84 mins

list.cod.decomp.tables.smoothed <- vector(mode = "list", length = dim(paired.ids)[1])

system.time(
for(i in 1:dim(paired.ids)[1]) {
  white.index <- which(non.zero.stratums.smoothed == paired.ids[i, "StateYearRaceSexWhite"])
  black.index <- which(non.zero.stratums.smoothed == paired.ids[i, "StateYearRaceSexBlack"])

  if(length(white.index) != 0 & length(black.index) != 0) {
    ct.white <- subset(dat.clean, StateYearRaceSex == paired.ids[i, "StateYearRaceSexWhite"]) %>% 
                ungroup() %>%
                select(Age2, COD2, cod_prop_smoothed_deaths) %>%
                dplyr::rename(prop.white = cod_prop_smoothed_deaths)

    ct.black <- subset(dat.clean, StateYearRaceSex == paired.ids[i, "StateYearRaceSexBlack"]) %>% 
                ungroup() %>%
                select(Age2, COD2, cod_prop_smoothed_deaths) %>%
                dplyr::rename(prop.black = cod_prop_smoothed_deaths)
    ct.both <- merge(ct.white, ct.black, by = c("Age2", "COD2")) 
    ct.both <- ct.both %>% arrange(COD2, Age2)

    if(any(is.na(ct.both$prop.black)) == F & any(is.na(ct.both$prop.white)) == F) { #ensures that there are no ages with 0 deaths - for these COD prop == 0 and can't compute

        list.cod.decomp.tables.smoothed[[i]] <- cause_of_death_decomp(life.table1 = as.data.frame(list.life.tables.smoothed[[black.index]]),
                                                                      life.table2 = as.data.frame(list.life.tables.smoothed[[white.index]]), 
                                                                      decomp.table = as.data.frame(list.age.decomp.tables.smoothed[[i]]),
                                                                      cod.table = ct.both, 
                                                                      age.colname.cod.table = "Age2", 
                                                                      COD.colname.cod.table = "COD2", 
                                                                      prop1.colname.cod.table = "prop.black", 
                                                                      prop2.colname.cod.table = "prop.white")

    } else { list.cod.decomp.tables.smoothed[[i]] <- "Encountered age group with no deaths"

    }
  } else {

    list.cod.decomp.tables.smoothed[[i]] <- "Couldn't compare life tables"
  }

  print(i)
}
)
```

```{r update_decomp_for_plotting}
#update the decomposition table for plotting purposes
for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables.smoothed[[i]][1] != "Couldn't compare life tables" & 
       list.cod.decomp.tables.smoothed[[i]][1] != "Encountered age group with no deaths") {

       list.cod.decomp.tables.smoothed[[i]] <- make_dataset_cod_plot(cod.decomp.table = list.cod.decomp.tables.smoothed[[i]],
                                                            age.groups = "Ages", 
                                                            cause.of.death = "Cause.of.death")  
    } 
  print(i)
} 
```

```{r compute_marginal_COD_decomposition}
list.cod.marginal.tables.smoothed <- vector(mode = "list", length = dim(paired.ids)[1])

for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables.smoothed[[i]][1] != "Couldn't compare life tables" & 
       list.cod.decomp.tables.smoothed[[i]][1] != "Encountered age group with no deaths") { 
       list.cod.marginal.tables.smoothed[[i]] <-  list.cod.decomp.tables.smoothed[[i]] %>%
                                                  dplyr::group_by(Cause.of.death) %>% 
                                                  dplyr::summarise(C_x_COD = sum(C_xi)) %>%
                                                  dplyr::mutate(C_x_COD_proportion = C_x_COD/sum(C_x_COD),
                                                         adds_to_gap = ifelse(C_x_COD > 0, T, F))

    }
}
```

```{r save_workspace}
save.image(file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/Sep6_BWgap.Rdata")
```

```{r save_specified_files}
save(dat.clean, dat2, BlackWhite, BlackWhiteSmooth, paired.ids, list.age.decomp.tables, list.cod.decomp.tables,
     list.age.decomp.tables5, list.cod.decomp.tables5, list.age.decomp.tables9, list.cod.decomp.tables9,
     list.age.decomp.tables.smoothed, list.cod.decomp.tables.smoothed, list.cod.marginal.tables.smoothed,
     non.zero.stratums, non.zero.stratums.smoothed,
     file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/main_datasets.Rdata")
```

Curious how this affects the LE calculation 
- using fractional data
- The fact that smoothed are NEVER zero seems problematic to me because it will bias LE downwards
- Perhaps that is not the case because it also pulls some death counts down.

- Best way to investigate this? Are there any life tables where we have 0 counts (exactly known) but no NA counts?
- need to do a summary with info on num.zeroes and num.NAs and see if any strata have num.zeroes > 1 but num.NA == 0.