### 2017-01-18 Code/Run_model_on_SH_data/Prep SH HA data for our analysis.Rmd

---
title: "Compare with SH HA paper"
author: "Corinne Riddell"
date: "December 13, 2016"
output: html_document
---

```{r}
library(foreign)
library(dplyr)

dtotal <- read.table("/Users/corinneriddell/Dropbox/BlackWhiteGap/dataverse_files/deaths-total-nhisp.txt", header = F, sep="\t")

names(dtotal)
#no names affixed to dataset. But the names are stored in the data dictionary file found at:
#/Users/corinneriddell/Dropbox/BlackWhiteGap/dataverse_files/deaths-total-nhisp.dic

str(dtotal)
table(dtotal$V1) #50 states + DC
table(dtotal$V2) #20 age groups
table(dtotal$V3) #sex
table(dtotal$V4) #race
table(dtotal$V5) #20 years

names(dtotal) <- c("State", "Age", "Sex", "Race", "Year", "Crude_Rate", "Count", "Population") #see the data dictionary - path is above.
```

```{r recode_variables}
dtotal <- dtotal %>%
  mutate(State2 = factor(State, levels = c(0:50), labels = c("Alabama", "Alaska", "Arizona", "Arkansas","California",
                                                             "Colorado", "Connecticut", "Delaware", "Washington DC",
                                                             "Florida", "Georgia", "Hawaii","Idaho", "Illinois", 
                                                             "Indiana","Iowa", "Kansas", "Kentucky", "Louisiana", 
                                                             "Maine", "Maryland", "Massachusetts", "Michigan", 
                                                             "Minnesota", "Mississippi","Missouri", "Montana", 
                                                             "Nebraska", "Nevada", "New Hampshire", "New Jersey",
                                                             "New Mexico", "New York","North Carolina","North Dakota",
                                                             "Ohio", "Oklahoma","Oregon", "Pennsylvania", 
                                                             "Rhode Island", "South Carolina", "South Dakota", 
                                                             "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
                                                             "Washington", "West Virginia", "Wisconsin", "Wyoming")), 
         Age2 = factor(Age, levels = c(0:19), labels = c("<1 year", "1-4 years", "5-9 years", "10-14 years", 
                                                         "15-19 years", "20-24 years", "25-29 years", "30-34 years",
                                                         "35-39 years", "40-44 years", "45-49 years", "50-54 years", 
                                                         "55-59 years", "60-64 years", "65-69 years", "70-74 years", 
                                                         "75-79 years", "80-84 years","85+ years", "Unknown"), ordered = T),
         Age3 = 0*(Age2 == "<1 year") + 1*(Age2 == "1-4 years") + 5*(Age2 == "5-9 years") +
                10*(Age2 == "10-14 years") + 15*(Age2 == "15-19 years") + 20*(Age2 == "20-24 years") + 
                25*(Age2 == "25-29 years") + 30*(Age2 == "30-34 years") + 35*(Age2 == "35-39 years") +
                40*(Age2 == "40-44 years") + 45*(Age2 == "45-49 years") + 50*(Age2 == "50-54 years") + 
                55*(Age2 == "55-59 years") + 60*(Age2 == "60-64 years") + 65*(Age2 == "65-69 years") +   
                70*(Age2 == "70-74 years") + 75*(Age2 == "75-79 years") + 80*(Age2 == "80-84 years") + 
                85*(Age2 == "85+ years"),
         nx = 1*(Age2 == "<1 year") + 4*(Age2 == "1-4 years") + 5*(Age2 != "<1 year" & Age2 != "1-4 years"),
         Sex2 = factor(Sex, levels = c(0, 1), labels = c("Male", "Female")),
         Race2 = factor(Race, levels = c(0, 1), labels = c("White", "Black")),
         Year2 = factor(Year, levels = c(0:19), labels = as.character(1990:2009)),
         Year3 = as.numeric(as.character(Year2)),
         RaceSex = interaction(Race2, Sex2, sep = " "),
         StateYearSex = interaction(State2, Year2, Sex2),
         StateYearRaceSex = interaction(State2, Year2, Race2, Sex2))
```

```{r unknown_age_group}
#there is an unknown age group for each combo of state-year-race-sex-age:
table(dtotal$Age2)

#Count of death is always equal to 0 for unknown age groups
table(dtotal$Age, dtotal$Count==0)

#remove these data from the data we analyze:
dtotal <- dtotal[dtotal$Age2 != "Unknown", ]

save(dtotal, file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/SH_HA_data.Rdata")
```