### 2017-01-20 Code/2_Run_LE_and_Decomp_Calcs.Rmd

---
title: "Run Life expectancy and decomposition calculations"
author: "Corinne Riddell"
date: "1/19/2017"
output: html_document
---

Depends on: output datasets created in "1_Explore_Data.Rmd".

```{r load_libraries, echo=F, warning=F, message=F}
library(tidyverse)
library(plotly)
options(tibble.width = Inf) #display all columns of tibbles
```

```{r load_my_functions}
source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")
```

```{r load_data}
dat.aggregated <- read.csv("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/dat_aggregated.Rdata", header = T)
dat.clean<- read.csv("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/dat_clean.Rdata", header = T)
```

In those strata that do not have any age-specific population counts of 0, create a life table and store it in a list.

```{r calculate_life_expectancy}
non.zero.stratums <- setdiff(levels(dat.aggregated$StateYearRaceSex), 
                             unique(dat.aggregated$StateYearRaceSex[dat.aggregated$Population == 0]))

#takes 13 mins

list.life.tables1 <- list.life.tables5 <- list.life.tables9 <- vector(mode = "list", length = length(non.zero.stratums))
list.index <- 1

system.time(
  for(stratum.id in non.zero.stratums) {

    list.life.tables1[[list.index]]$stratum.id <- non.zero.stratums[list.index]
    list.life.tables5[[list.index]]$stratum.id <- non.zero.stratums[list.index]
    list.life.tables9[[list.index]]$stratum.id <- non.zero.stratums[list.index]

    list.life.tables1[[list.index]]$life.table <- life.table(data = subset(dat.aggregated, StateYearRaceSex == stratum.id) %>% arrange(Age_factor),
                                                 num.ages.in.group = "Age_bin_size",
                                                 death.counts = "total_deaths1",
                                                 population.counts = "Population")

    list.life.tables5[[list.index]]$life.table <- life.table(data = subset(dat.aggregated, StateYearRaceSex == stratum.id) %>% arrange(Age_factor),
                                                  num.ages.in.group = "Age_bin_size",
                                                  death.counts = "total_deaths5",
                                                  population.counts = "Population")

    list.life.tables9[[list.index]]$life.table <- life.table(data = subset(dat.aggregated, StateYearRaceSex == stratum.id) %>% arrange(Age_factor),
                                                  num.ages.in.group = "Age_bin_size",
                                                  death.counts = "total_deaths9",
                                                  population.counts = "Population")
    list.index <- list.index + 1
    print(list.index)
  }
)
```

Add the life expectancy at birth to the data file:

```{r add_le_to_dataset}
le1 <- map_dbl(list.life.tables1, ~ .x$life.table$e_x[1])
le5 <- map_dbl(list.life.tables5, ~ .x$life.table$e_x[1])
le9 <- map_dbl(list.life.tables9, ~ .x$life.table$e_x[1])

sid <- map_chr(list.life.tables1, ~ .x$stratum.id)

les <- tibble(le1, le5, le9, sid)
names(les) = c("life.expectancy.birth.1", "life.expectancy.birth.5", "life.expectancy.birth.9", "StateYearRaceSex")

dat.aggregated <- merge(dat.aggregated, les, by = "StateYearRaceSex", all.x =T)

rm(les, le1, le5, le9)
```

```{r densityplot_life_expectancy}
ggplot(dat.aggregated %>% dplyr::filter(Age_factor == "<1 year"), aes(x = life.expectancy.birth.1)) + 
  geom_density(aes(fill="Impute 1"), alpha = 0.3) +
  geom_density(aes(x = life.expectancy.birth.5, fill = "Impute 5"), alpha = 0.3) +
  geom_density(aes(x = life.expectancy.birth.9, fill="Impute 9"), alpha = 0.3) +
  facet_grid(Sex_factor ~ Race_factor)
```

Make a new dataset that contains information on the gap:

```{r create_dataset_BlackWhite}
#this dataset contains the black-white data within the same row (wide format) for each year-state-sex strata

Black <- dat.aggregated %>% filter(Race_factor == "Black" & Age_minbin == 0) %>% 
                            select(State_factor, Year_actual, Sex_factor, Pop_across_age, 
                                   life.expectancy.birth.1, life.expectancy.birth.5, 
                                   life.expectancy.birth.9) %>%
                            rename(Pop_black = Pop_across_age,  
                                   le.birth.black1 = life.expectancy.birth.1, 
                                   le.birth.black5 = life.expectancy.birth.5, 
                                   le.birth.black9 = life.expectancy.birth.9)

White <- dat.aggregated %>% filter(Race_factor == "White" & Age_minbin == 0) %>% 
                            select(State_factor, Year_actual, Sex_factor, Pop_across_age, 
                                   life.expectancy.birth.1, life.expectancy.birth.5, 
                                   life.expectancy.birth.9) %>%
                            rename(Pop_white = Pop_across_age,  
                                   le.birth.white1 = life.expectancy.birth.1,
                                   le.birth.white5 = life.expectancy.birth.5,
                                   le.birth.white9 = life.expectancy.birth.9)

BlackWhite <- merge(Black, White, by = c("State_factor", "Year_actual", "Sex_factor"))

BlackWhite <- BlackWhite %>% mutate(WBgap1 = le.birth.white1 - le.birth.black1,
                                    WBgap5 = le.birth.white5 - le.birth.black5,
                                    WBgap9 = le.birth.white9 - le.birth.black9)

rm(Black, White)
```

Decompose the gap by age. That is, for each year-state-sex combination, compare the life table for blacks to whites and decompose the gap in life expectancy at birth by age.

```{r create_paired_ids_dataset}
#paired.ids will contain the index names that are sent to the age decomposition function in the next code chunk.
#it pairs ids between strata for blacks and whites

#first, create list of stratum.ids for each race-sex combination
for (rs in unique(levels(dat.clean$RaceSex))) {
  assign(paste0(gsub(" ", "", rs), ".stratum.ids"), 
         dat.clean %>% 
           ungroup() %>%
           filter(RaceSex == rs) %>%
           select(StateYearRaceSex, StateYearSex, Sex_factor, State_factor, Year_actual) %>%
           group_by(StateYearRaceSex, StateYearSex) %>%
           filter(row_number() == 1)
  )

}

#then, merge the black-white lists
Female.ids <- merge(WhiteFemale.stratum.ids, BlackFemale.stratum.ids[, -c(3:5)], 
                    by = "StateYearSex", suffixes = c("White", "Black"))
head(Female.ids)

Male.ids <- merge(WhiteMale.stratum.ids, BlackMale.stratum.ids[, -c(3:5)],
                  by = "StateYearSex", suffixes = c("White", "Black"))
head(Male.ids)

paired.ids <- rbind(Male.ids, Female.ids)

rm(WhiteFemale.stratum.ids, BlackFemale.stratum.ids, WhiteMale.stratum.ids, BlackMale.stratum.ids, Female.ids, Male.ids)
```

```{r calculate_age_decomp}
#runs in about a minute

list.age.decomp.tables1 <- list.age.decomp.tables5 <- list.age.decomp.tables9 <- vector(mode = "list", length = dim(paired.ids)[1])

system.time(
  for(i in 1:dim(paired.ids)[1]) {
    white.index <- which(non.zero.stratums == paired.ids[i, "StateYearRaceSexWhite"])
    black.index <- which(non.zero.stratums == paired.ids[i, "StateYearRaceSexBlack"])

    list.age.decomp.tables1[[i]]$StateYearSex <- paired.ids[i, "StateYearSex"]
    list.age.decomp.tables5[[i]]$StateYearSex <- paired.ids[i, "StateYearSex"]
    list.age.decomp.tables9[[i]]$StateYearSex <- paired.ids[i, "StateYearSex"]

    if(length(white.index) != 0 & length(black.index) != 0) {
      list.age.decomp.tables1[[i]]$decomp.table <- le_age_decomp(life.table1 = as.data.frame(list.life.tables1[[black.index]]$life.table), 
                                                                 name.lt1 = "Black", 
                                                                 life.table2 = as.data.frame(list.life.tables1[[white.index]]$life.table), 
                                                                 name.lt2 = "White", age.groups = "Age_factor")

      list.age.decomp.tables5[[i]]$decomp.table <- le_age_decomp(life.table1 = as.data.frame(list.life.tables5[[black.index]]$life.table), 
                                                                 name.lt1 = "Black", 
                                                                 life.table2 = as.data.frame(list.life.tables5[[white.index]]$life.table), 
                                                                 name.lt2 = "White", age.groups = "Age_factor")

      list.age.decomp.tables9[[i]]$decomp.table <- le_age_decomp(life.table1 = as.data.frame(list.life.tables9[[black.index]]$life.table), 
                                                                 name.lt1 = "Black", 
                                                                 life.table2 = as.data.frame(list.life.tables9[[white.index]]$life.table), 
                                                                 name.lt2 = "White", age.groups = "Age_factor")
    } else {
      list.age.decomp.tables1[[i]]$decomp.table <- list.age.decomp.tables5[[i]]$decomp.table <- list.age.decomp.tables9[[i]]$decomp.table <-"Couldn't compare life tables"
    }
  }
)

```

```{r save_objects}
write_csv(dat.aggregated, path = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/dat_aggregated2.csv", col_names = T)
write_csv(BlackWhite, path = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/BlackWhite.csv", col_names = T)

save(paired.ids, non.zero.stratums,
     list.age.decomp.tables1, list.age.decomp.tables5, list.age.decomp.tables9,
     list.life.tables1, list.life.tables5, list.life.tables9,  
     file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/calculated_data.Rdata")
```

```{r calculate_cod_decomp}
#stopped here - need to figure this out
#if the count of deaths == 0 for all causes of death within an age band then the COD proportion is NA
#this messes up the cod decomposition. What to do?
#example - no black people dying between the ages of 5-9 of anything. but have whites dying in that age group so their
#COD proportions are okay.

#takes 1.5 hours for all three
list.cod.decomp.tables1 <- list.cod.decomp.tables5 <- list.cod.decomp.tables9 <- vector(mode = "list", length = dim(paired.ids)[1])

system.time(
  for(i in 1:dim(paired.ids)[1]) {
    white.index <- which(non.zero.stratums == paired.ids[i, "StateYearRaceSexWhite"])
    black.index <- which(non.zero.stratums == paired.ids[i, "StateYearRaceSexBlack"])

    list.cod.decomp.tables1[[i]]$StateYearSex <- paired.ids[i, "StateYearSex"]
    list.cod.decomp.tables5[[i]]$StateYearSex <- paired.ids[i, "StateYearSex"]
    list.cod.decomp.tables9[[i]]$StateYearSex <- paired.ids[i, "StateYearSex"]

    if(length(white.index) != 0 & length(black.index) != 0) {
      ct.white <- subset(dat.clean, StateYearRaceSex == paired.ids[i, "StateYearRaceSexWhite"]) %>% 
        ungroup() %>%
        select(Age_factor, COD_factor, cod_prop_death1, cod_prop_death5, cod_prop_death9) %>%
        rename(prop.white1 = cod_prop_death1, prop.white5 = cod_prop_death5, prop.white9 = cod_prop_death9)

      ct.black <- subset(dat.clean, StateYearRaceSex == paired.ids[i, "StateYearRaceSexBlack"]) %>% 
        ungroup() %>%
        select(Age_factor, COD_factor, cod_prop_death1, cod_prop_death5, cod_prop_death9) %>%
        rename(prop.black1 = cod_prop_death1, prop.black5 = cod_prop_death5, prop.black9 = cod_prop_death9)

      ct.both <- merge(ct.white, ct.black, by = c("Age_factor", "COD_factor")) 
      ct.both <- ct.both %>% arrange(COD_factor, Age_factor)

      if(!any(is.na(ct.both$prop.black1)) & !any(is.na(ct.both$prop.white1))) { #ensures that there are no ages with 0 deaths - for these COD prop == 0 and can't compute

        list.cod.decomp.tables1[[i]]$decomp.table <- cause_of_death_decomp(life.table1 = as.data.frame(list.life.tables1[[black.index]]$life.table),
                                                              life.table2 = as.data.frame(list.life.tables1[[white.index]]$life.table), 
                                                              decomp.table = as.data.frame(list.age.decomp.tables1[[i]]$decomp.table),
                                                              cod.table = ct.both %>% select(Age_factor, COD_factor, prop.white1, prop.black1), 
                                                              age.colname.cod.table = "Age_factor", 
                                                              COD.colname.cod.table = "COD_factor", 
                                                              prop1.colname.cod.table = "prop.black1", 
                                                              prop2.colname.cod.table = "prop.white1")

        list.cod.decomp.tables5[[i]]$decomp.table <- cause_of_death_decomp(life.table1 = as.data.frame(list.life.tables5[[black.index]]$life.table),
                                                              life.table2 = as.data.frame(list.life.tables5[[white.index]]$life.table), 
                                                              decomp.table = as.data.frame(list.age.decomp.tables5[[i]]$decomp.table),
                                                              cod.table = ct.both %>% select(Age_factor, COD_factor, prop.white5, prop.black5), 
                                                              age.colname.cod.table = "Age_factor", 
                                                              COD.colname.cod.table = "COD_factor", 
                                                              prop1.colname.cod.table = "prop.black5", 
                                                              prop2.colname.cod.table = "prop.white5")

        list.cod.decomp.tables9[[i]]$decomp.table <- cause_of_death_decomp(life.table1 = as.data.frame(list.life.tables5[[black.index]]$life.table),
                                                              life.table2 = as.data.frame(list.life.tables5[[white.index]]$life.table), 
                                                              decomp.table = as.data.frame(list.age.decomp.tables5[[i]]$decomp.table),
                                                              cod.table = ct.both %>% select(Age_factor, COD_factor, prop.white9, prop.black9), 
                                                              age.colname.cod.table = "Age_factor", 
                                                              COD.colname.cod.table = "COD_factor", 
                                                              prop1.colname.cod.table = "prop.black9", 
                                                              prop2.colname.cod.table = "prop.white9")

      } else { list.cod.decomp.tables1[[i]]$decomp.table <- list.cod.decomp.tables5[[i]]$decomp.table <- list.cod.decomp.tables9[[i]]$decomp.table <- "Encountered age group with no deaths"

      }
    } else {

      list.cod.decomp.tables1[[i]]$decomp.table <- list.cod.decomp.tables5[[i]]$decomp.table <- list.cod.decomp.tables9[[i]]$decomp.table <- "Couldn't compare life tables"
    }

    print(i)
  }
)
```

```{r  update_decomp_for_plotting}
#update the decomposition table for plotting purposes
for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables[[i]][1] != "Couldn't compare life tables" & 
       list.cod.decomp.tables[[i]][1] != "Encountered age group with no deaths") {

       list.cod.decomp.tables[[i]] <- make_dataset_cod_plot(cod.decomp.table = list.cod.decomp.tables1[[i]],
                                                            age.groups = "Ages", 
                                                            cause.of.death = "Cause.of.death", 
                                                            sign.var = "sign", decomp.var = "C_xi", 
                                                            decomp.var.prop = "C_xi_proportion")

       list.cod.decomp.tables5[[i]] <- make_dataset_cod_plot(cod.decomp.table = list.cod.decomp.tables5[[i]],
                                                            age.groups = "Ages", 
                                                            cause.of.death = "Cause.of.death", 
                                                            sign.var = "sign", decomp.var = "C_xi", 
                                                            decomp.var.prop = "C_xi_proportion")

       list.cod.decomp.tables9[[i]] <- make_dataset_cod_plot(cod.decomp.table = list.cod.decomp.tables9[[i]],
                                                            age.groups = "Ages", 
                                                            cause.of.death = "Cause.of.death", 
                                                            sign.var = "sign", decomp.var = "C_xi", 
                                                            decomp.var.prop = "C_xi_proportion") 
    } 
  print(i)
}  
```

```{r pretty_cod}
ggplotly(ggplot(list.cod.decomp.tables[[1]], aes(y = Ages, x = start)) + 
           geom_segment(aes(xend = finish, col = Cause.of.death, yend = Ages), lwd = 4) +
           xlab("Contribution to life expectancy gap") +
           ylab("Age group\n") + theme_minimal() +
           geom_vline(xintercept = 0))

ggplotly(ggplot(list.cod.decomp.tables9[[1]], aes(y = Ages, x = start)) + 
           geom_segment(aes(xend = finish, col = Cause.of.death, yend = Ages), lwd = 4) +
           xlab("Contribution to life expectancy gap") +
           ylab("Age group\n") + theme_minimal() +
           geom_vline(xintercept = 0))
```