### 2017-01-27 Code/3_validation/run-models.Rmd
alabama males results!

---
title: "Run models"
output: github_document
---

```{r load_libraries}
library(Rcpp)
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(rjags)
library(R2jags) 
library(devtools)
library(jagstools)
library(reshape2)
library(parallel)
library(gtools)
library(bayesboot)
library(coda)
library(grid)
library(gridExtra)
```

```{r load_user_written_functions}
source("~/BlackWhiteMortalityGap/Code/2_smoothing_models/bayes_smoothing_functions.R")
source('~/BlackWhiteMortalityGap/Code/life_expectancy_functions.R') 
```

```{r load_tidied_data}
dat.clean = read.csv('~/dat_clean.csv')
```

```{r fit_models_alabama_males}
#run time: 15 minutes
#fit the bayesian models and put the 1000 posterior samples into a data frame
r.White <- run_analysis(dataset = dat.clean, state1 = "Alabama", sex1 = "Male", race1 = "White")
r.White$race <- "White"

r.Black <- run_analysis(dataset = dat.clean, state1 = "Alabama", sex1 = "Male", race1 = "Black")
r.Black$race <- "Black"
```

```{r prepare_data_for_le_functions}
#run time: about 2 minutes
r.All <- rbind(r.White, r.Black)
rm(r.White, r.Black)

r.All$smoothed_deaths <- r.All$smoothed_rate*r.All$population

r.All <- r.All %>% group_by(post.samp, state, race.sex, year.n, age_minbin) %>%
                           arrange(post.samp, state, race.sex, year.n, age_minbin) %>%
                           mutate(total_smoothed_deaths = sum(smoothed_deaths),
                                  cod_prop_smoothed_deaths = smoothed_deaths/total_smoothed_deaths)
```

```{r run_calculations}
#run time ~40 mins
#run the life table and decomposition calculations for each of the 1000 posterior samples

n <- length(levels(droplevels(r.All$state.year.sex)))

le.calculations = vector(mode = "list", length = n)
counter = 1
system.time(
  for (stratum_i in levels(droplevels(r.All$state.year.sex))){

    data_sub <- subset(r.All, state.year.sex == stratum_i)

    le.calculations[[counter]]$stratum.id <- stratum_i
    le.calculations[[counter]]$calcs <- by(data = data_sub,
                                           INDICES = list(data_sub$post.samp),
                                           FUN = function(x) life_expectancy_and_gap(data = x))

    counter <- counter + 1

    print(stratum_i)
    rm(data_sub)
  }
)
rm(counter)

```

```{r extract_life_expectancy_results}
BlackWhite <- data.frame(stratum.id = character(),
                         LE_white_lcl = numeric(), LE_white = numeric(), LE_white_ucl = numeric(),
                         LE_black_lcl = numeric(), LE_black = numeric(), LE_black_ucl = numeric(),
                         LE_wbgap_lcl = numeric(), LE_wbgap = numeric(), LE_wbgap_ucl = numeric(),
                         stringsAsFactors = F)

age_minbin <- le.calculations[[1]]$calcs[[1]]$age.decomp$Ages
cods <- le.calculations[[1]]$calcs[[1]]$cod.marginal$Cause.of.death

age.decomp.estimates <- data.frame(stratum.id = rep(levels(droplevels(r.All$state.year.sex)), 
                                                    each = length(age_minbin)),
                                   age_minbin = numeric(19*n),
                                   age_cont_yrs_lcl = numeric(19*n), 
                                   age_cont_yrs = numeric(19*n), 
                                   age_cont_yrs_ucl = numeric(19*n),
                                   age_cont_prop_lcl = numeric(19*n), 
                                   age_cont_prop = numeric(19*n), 
                                   age_cont_prop_ucl = numeric(19*n))

cod.marginal.estimates <- data.frame(stratum.id = rep(levels(droplevels(r.All$state.year.sex)), 
                                                    each = length(cods)),
                                   COD = character(6*n),
                                   COD_cont_yrs_lcl = numeric(6*n), 
                                   COD_cont_yrs = numeric(6*n), 
                                   COD_cont_yrs_ucl = numeric(6*n),
                                   COD_cont_prop_lcl = numeric(6*n), 
                                   COD_cont_prop = numeric(6*n), 
                                   COD_cont_prop_ucl = numeric(6*n), 
                                   stringsAsFactors = F)

age.cod.estimates <- data.frame(stratum.id = rep(levels(droplevels(r.All$state.year.sex)), 
                                                 each = length(cods)*length(age_minbin)),
                                age_minbin = numeric(19*6*n),   
                                COD = character(19*6*n),
                                age_COD_cont_yrs_lcl = numeric(19*6*n), 
                                age_COD_cont_yrs = numeric(19*6*n), 
                                age_COD_cont_yrs_ucl = numeric(19*6*n),
                                age_COD_cont_prop_lcl = numeric(19*6*n), 
                                age_COD_cont_prop = numeric(19*6*n), 
                                age_COD_cont_prop_ucl = numeric(19*6*n), 
                                stringsAsFactors = F)

for(i in 1:n) {

  BlackWhite[i, c("stratum.id")] <- le.calculations[[i]]$stratum.id

  #LIFE EXPECTANCY EXTRACT RESULTS
  BlackWhite[i , c("LE_white_lcl", "LE_white", "LE_white_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                    ~.x$le.df$LE_White), 
                                                                            c(0.025, 0.5, 0.975))
  BlackWhite[i , c("LE_black_lcl", "LE_black", "LE_black_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                    ~.x$le.df$LE_Black), 
                                                                            c(0.025, 0.5, 0.975))
  BlackWhite[i , c("LE_wbgap_lcl", "LE_wbgap", "LE_wbgap_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                    ~.x$le.df$LE_WBgap), 
                                                                            c(0.025, 0.5, 0.975))

  #AGE DECOMPOSITION EXTRACT RESULTS
  age.decomp.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$age.decomp$C_x)
  age.decomp.posteriors$age_minbin <- age_minbin
  age.decomp.posteriors <- age.decomp.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x")

  age.decomp.estimates[age.decomp.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                       c("age_minbin", "age_cont_yrs_lcl", "age_cont_yrs", "age_cont_yrs_ucl")] <- 
    age.decomp.posteriors %>% 
    group_by(age_minbin) %>%
    summarise(lcl_C_x = quantile(C_x, 0.025),
              median_C_x = quantile(C_x, 0.5), 
              ucl_C_x = quantile(C_x, 0.975))

  rm(age.decomp.posteriors)

  age.decomp.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$age.decomp$C_x_proportion)
  age.decomp.posteriors2$age_minbin <- age_minbin
  age.decomp.posteriors2 <- age.decomp.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x_proportion")

  age.decomp.estimates[age.decomp.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                       c("age_minbin", "age_cont_prop_lcl", "age_cont_prop", "age_cont_prop_ucl")] <- 
    age.decomp.posteriors2 %>% 
    group_by(age_minbin) %>%
    summarise(lcl_C_x_proportion = quantile(C_x_proportion, 0.025),
              median_C_x_proportion = quantile(C_x_proportion, 0.5), 
              ucl_C_x_proportion = quantile(C_x_proportion, 0.975))

  rm(age.decomp.posteriors2)

  #COD DECOMPOSITION EXTRACT RESULTS
  cod.marginal.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.marginal$C_x)
  cod.marginal.posteriors$COD <- as.character(cods)
  cod.marginal.posteriors <- cod.marginal.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x")

  cod.marginal.estimates[cod.marginal.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                       c("COD", "COD_cont_yrs_lcl", "COD_cont_yrs", "COD_cont_yrs_ucl")] <- 
    cod.marginal.posteriors %>% 
    group_by(COD) %>%
    summarise(lcl_C_x = quantile(C_x, 0.025),
              median_C_x = quantile(C_x, 0.5), 
              ucl_C_x = quantile(C_x, 0.975))

  rm(cod.marginal.posteriors)

  cod.marginal.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.marginal$C_x_proportion)
  cod.marginal.posteriors2$COD <- as.character(cods)
  cod.marginal.posteriors2 <- cod.marginal.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", 
                                                                         value = "C_x_proportion")

  cod.marginal.estimates[cod.marginal.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                       c("COD", "COD_cont_prop_lcl", "COD_cont_prop", "COD_cont_prop_ucl")] <-
    cod.marginal.posteriors2 %>%
    group_by(COD) %>%
    summarise(lcl_C_x_proportion = quantile(C_x_proportion, 0.025),
              median_C_x_proportion = quantile(C_x_proportion, 0.5),
              ucl_C_x_proportion = quantile(C_x_proportion, 0.975))

  rm(cod.marginal.posteriors)

  #AGE-COD EXTRACT RESULTS
  age.cod.decomp.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.decomp$C_xi)
  age.cod.decomp.posteriors$age_minbin <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Ages
  age.cod.decomp.posteriors$COD <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Cause.of.death
  age.cod.decomp.posteriors <- age.cod.decomp.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_xi")

  age.cod.estimates[age.cod.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                       c("age_minbin", "COD", "age_COD_cont_yrs_lcl", "age_COD_cont_yrs", "age_COD_cont_yrs_ucl")] <-
    age.cod.decomp.posteriors %>%
    group_by(age_minbin, COD) %>%
    summarise(lcl_C_x = quantile(C_xi, 0.025),
              median_C_x = quantile(C_xi, 0.5),
              ucl_C_x = quantile(C_xi, 0.975))

  rm(age.cod.decomp.posteriors)

  age.cod.decomp.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.decomp$C_xi_proportion)
  age.cod.decomp.posteriors2$age_minbin <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Ages
  age.cod.decomp.posteriors2$COD <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Cause.of.death
  age.cod.decomp.posteriors2 <- age.cod.decomp.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", 
                                                                             value = "C_xi_proportion")

  age.cod.estimates[age.cod.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                       c("age_minbin", "COD", "age_COD_cont_prop_lcl", "age_COD_cont_prop", "age_COD_cont_prop_ucl")] <-
    age.cod.decomp.posteriors2 %>%
    group_by(age_minbin, COD) %>%
    summarise(lcl_C_x_proportion = quantile(C_xi_proportion, 0.025),
              median_C_x_proportion = quantile(C_xi_proportion, 0.5),
              ucl_C_x_proportion = quantile(C_xi_proportion, 0.975))

  rm(age.cod.decomp.posteriors2)
}
```

```{r}
# load('~/SH_HA_data.Rdata')
# SH_sub <- dtotal %>% filter(state == "Alabama", sex == "Male", race == "Black", year == 1990)
# 
# SH_sub %>% select(age, age_minbin, deaths, population)
```