### 2017-01-30 Code/3_validation/run-models.Rmd
models running

---
title: "Run models"
output: github_document
---

```{r load_libraries}
library(Rcpp)
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(rjags)
library(R2jags) 
library(devtools)
library(jagstools)
library(reshape2)
library(parallel)
library(gtools)
library(bayesboot)
library(coda)
library(grid)
library(gridExtra)
library(parallel)
```

```{r load_user_written_functions}
source("~/BlackWhiteMortalityGap/Code/2_smoothing_models/bayes_smoothing_functions.R")
source('~/BlackWhiteMortalityGap/Code/life_expectancy_functions.R') 
```

```{r load_tidied_data}
dat.clean = read.csv('~/dat_clean.csv') 
```

```{r fit_models_alabama_males}
#run time: expect 2 hours
#fit the bayesian models and put the 1000 posterior samples into a data frame

list_set2 <- list("Colorado", "Connecticut", "Delaware", "Washington DC", "Florida", "Georgia") #saved as results_set_2

system.time(
  for(sex_i in c("Male", "Female")){
    #for(state_i in c("Alabama", "Arizona", "Arkansas", "California")) { #saved as results_set_1
    for(state_i in c("Colorado", "Connecticut", "Delaware", "Washington DC")) { #saved as results_set_2
      assign(paste0("r.White.", gsub(" ", "", state_i), ".", sex_i), run_analysis(dataset = dat.clean, state1 = state_i, sex1 = sex_i, race1 = "White"))
      assign(paste0("r.Black.", gsub(" ", "", state_i), ".", sex_i), run_analysis(dataset = dat.clean, state1 = state_i, sex1 = sex_i, race1 = "Black"))
    }
  }
)

# r.White <- rbind(r.White.AlabamaMale, r.White.ArizonaMale, r.White.ArkansasMale, r.White.CaliforniaMale,
#                  r.White.AlabamaFemale, r.White.ArizonaFemale, r.White.ArkansasFemale, r.White.CaliforniaFemale)
r.White <- rbind(r.White.Colorado.Male, r.White.Connecticut.Male, r.White.Delaware.Male, r.White.WashingtonDC.Male,
                 r.White.Colorado.Female, r.White.Connecticut.Female, r.White.Delaware.Female, r.White.WashingtonDC.Female)
r.White$race <- "White"

# r.Black <- rbind(r.Black.AlabamaMale, r.Black.ArizonaMale, r.Black.ArkansasMale, r.Black.CaliforniaMale,
#                  r.Black.AlabamaFemale, r.Black.ArizonaFemale, r.Black.ArkansasFemale, r.Black.CaliforniaFemale)
r.White <- rbind(r.Black.Colorado.Male, r.Black.Connecticut.Male, r.Black.Delaware.Male, r.Black.WashingtonDC.Male,
                 r.Black.Colorado.Female, r.Black.Connecticut.Female, r.Black.Delaware.Female, r.Black.WashingtonDC.Female)
r.Black$race <- "Black"
```

```{r prepare_data_for_le_functions}
#run time: used to be 2 minutes
r.All <- rbind(r.White, r.Black)
rm(r.White, r.Black)

r.All$smoothed_deaths <- r.All$smoothed_rate*r.All$population
table(r.All$smoothed_deaths == 0) #check this -- if no deaths in an age-band (i.e., no death for any COD, then can't run the COD calculations for that stratum.)

system.time(r.All <- r.All %>% group_by(post.samp, state, race.sex, year.n, age_minbin) %>%
                           arrange(post.samp, state, race.sex, year.n, age_minbin))
```

```{r run_calculations}
#run time: 8.5 hours for four states
#run the life table and decomposition calculations for each of the 1000 posterior samples

n <- length(levels(droplevels(r.All$state.year.sex)))

le.calculations = vector(mode = "list", length = n)
counter = 1
system.time(
  for (stratum_i in levels(droplevels(r.All$state.year.sex))){

    data_sub <- subset(r.All, state.year.sex == stratum_i)

    le.calculations[[counter]]$stratum.id <- stratum_i
    le.calculations[[counter]]$calcs <- by(data = data_sub,
                                           INDICES = list(data_sub$post.samp),
                                           FUN = function(x) life_expectancy_and_gap(data = x))

    counter <- counter + 1

    print(stratum_i)
    rm(data_sub)
  }
)

```

```{r extract_life_expectancy_results}
BlackWhite <- data.frame(stratum.id = character(),
                         LE_white_lcl = numeric(), LE_white = numeric(), LE_white_ucl = numeric(),
                         LE_black_lcl = numeric(), LE_black = numeric(), LE_black_ucl = numeric(),
                         LE_wbgap_lcl = numeric(), LE_wbgap = numeric(), LE_wbgap_ucl = numeric(),
                         stringsAsFactors = F)

age_minbin <- le.calculations[[1]]$calcs[[1]]$age.decomp$Ages
cods <- le.calculations[[1]]$calcs[[1]]$cod.marginal$Cause.of.death

age.decomp.estimates <- data.frame(stratum.id = rep(levels(droplevels(r.All$state.year.sex)), 
                                                    each = length(age_minbin)),
                                   age_minbin = numeric(19*n),
                                   age_cont_yrs_lcl = numeric(19*n), 
                                   age_cont_yrs = numeric(19*n), 
                                   age_cont_yrs_ucl = numeric(19*n),
                                   age_cont_prop_lcl = numeric(19*n), 
                                   age_cont_prop = numeric(19*n), 
                                   age_cont_prop_ucl = numeric(19*n))

cod.marginal.estimates <- data.frame(stratum.id = rep(levels(droplevels(r.All$state.year.sex)), 
                                                    each = length(cods)),
                                   COD = character(6*n),
                                   COD_cont_yrs_lcl = numeric(6*n), 
                                   COD_cont_yrs = numeric(6*n), 
                                   COD_cont_yrs_ucl = numeric(6*n),
                                   COD_cont_prop_lcl = numeric(6*n), 
                                   COD_cont_prop = numeric(6*n), 
                                   COD_cont_prop_ucl = numeric(6*n), 
                                   stringsAsFactors = F)

age.cod.estimates <- data.frame(stratum.id = rep(levels(droplevels(r.All$state.year.sex)), 
                                                 each = length(cods)*length(age_minbin)),
                                age_minbin = numeric(19*6*n),   
                                COD = character(19*6*n),
                                age_COD_cont_yrs_lcl = numeric(19*6*n), 
                                age_COD_cont_yrs = numeric(19*6*n), 
                                age_COD_cont_yrs_ucl = numeric(19*6*n),
                                age_COD_cont_prop_lcl = numeric(19*6*n), 
                                age_COD_cont_prop = numeric(19*6*n), 
                                age_COD_cont_prop_ucl = numeric(19*6*n), 
                                stringsAsFactors = F)

system.time(
  for(i in 1:n) {

    BlackWhite[i, c("stratum.id")] <- le.calculations[[i]]$stratum.id

    #LIFE EXPECTANCY EXTRACT RESULTS
    BlackWhite[i , c("LE_white_lcl", "LE_white", "LE_white_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                      ~.x$le.df$LE_White), 
                                                                              c(0.025, 0.5, 0.975))
    BlackWhite[i , c("LE_black_lcl", "LE_black", "LE_black_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                      ~.x$le.df$LE_Black), 
                                                                              c(0.025, 0.5, 0.975))
    BlackWhite[i , c("LE_wbgap_lcl", "LE_wbgap", "LE_wbgap_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                      ~.x$le.df$LE_WBgap), 
                                                                              c(0.025, 0.5, 0.975))

    #AGE DECOMPOSITION EXTRACT RESULTS
    age.decomp.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$age.decomp$C_x)
    age.decomp.posteriors$age_minbin <- age_minbin
    age.decomp.posteriors <- age.decomp.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x")

    age.decomp.estimates[age.decomp.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                         c("age_minbin", "age_cont_yrs_lcl", "age_cont_yrs", "age_cont_yrs_ucl")] <- 
      age.decomp.posteriors %>% 
      group_by(age_minbin) %>%
      summarise(lcl_C_x = quantile(C_x, 0.025),
                median_C_x = quantile(C_x, 0.5), 
                ucl_C_x = quantile(C_x, 0.975))

    rm(age.decomp.posteriors)

    age.decomp.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$age.decomp$C_x_proportion)
    age.decomp.posteriors2$age_minbin <- age_minbin
    age.decomp.posteriors2 <- age.decomp.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x_proportion")

    age.decomp.estimates[age.decomp.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                         c("age_minbin", "age_cont_prop_lcl", "age_cont_prop", "age_cont_prop_ucl")] <- 
      age.decomp.posteriors2 %>% 
      group_by(age_minbin) %>%
      summarise(lcl_C_x_proportion = quantile(C_x_proportion, 0.025),
                median_C_x_proportion = quantile(C_x_proportion, 0.5), 
                ucl_C_x_proportion = quantile(C_x_proportion, 0.975))

    rm(age.decomp.posteriors2)

    #COD DECOMPOSITION EXTRACT RESULTS
    cod.marginal.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.marginal$C_x)
    cod.marginal.posteriors$COD <- as.character(cods)
    cod.marginal.posteriors <- cod.marginal.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x")

    cod.marginal.estimates[cod.marginal.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                           c("COD", "COD_cont_yrs_lcl", "COD_cont_yrs", "COD_cont_yrs_ucl")] <- 
      cod.marginal.posteriors %>% 
      group_by(COD) %>%
      summarise(lcl_C_x = quantile(C_x, 0.025),
                median_C_x = quantile(C_x, 0.5), 
                ucl_C_x = quantile(C_x, 0.975))

    rm(cod.marginal.posteriors)

    cod.marginal.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.marginal$C_x_proportion)
    cod.marginal.posteriors2$COD <- as.character(cods)
    cod.marginal.posteriors2 <- cod.marginal.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", 
                                                                           value = "C_x_proportion")

    cod.marginal.estimates[cod.marginal.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                           c("COD", "COD_cont_prop_lcl", "COD_cont_prop", "COD_cont_prop_ucl")] <-
      cod.marginal.posteriors2 %>%
      group_by(COD) %>%
      summarise(lcl_C_x_proportion = quantile(C_x_proportion, 0.025),
                median_C_x_proportion = quantile(C_x_proportion, 0.5),
                ucl_C_x_proportion = quantile(C_x_proportion, 0.975))

    rm(cod.marginal.posteriors)

    #AGE-COD EXTRACT RESULTS
    age.cod.decomp.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.decomp$C_xi)
    age.cod.decomp.posteriors$age_minbin <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Ages
    age.cod.decomp.posteriors$COD <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Cause.of.death
    age.cod.decomp.posteriors <- age.cod.decomp.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_xi")

    age.cod.estimates[age.cod.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                      c("age_minbin", "COD", "age_COD_cont_yrs_lcl", "age_COD_cont_yrs", "age_COD_cont_yrs_ucl")] <-
      age.cod.decomp.posteriors %>%
      group_by(age_minbin, COD) %>%
      summarise(lcl_C_x = quantile(C_xi, 0.025),
                median_C_x = quantile(C_xi, 0.5),
                ucl_C_x = quantile(C_xi, 0.975))

    rm(age.cod.decomp.posteriors)

    age.cod.decomp.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.decomp$C_xi_proportion)
    age.cod.decomp.posteriors2$age_minbin <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Ages
    age.cod.decomp.posteriors2$COD <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Cause.of.death
    age.cod.decomp.posteriors2 <- age.cod.decomp.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", 
                                                                               value = "C_xi_proportion")

    age.cod.estimates[age.cod.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                      c("age_minbin", "COD", "age_COD_cont_prop_lcl", "age_COD_cont_prop", "age_COD_cont_prop_ucl")] <-
      age.cod.decomp.posteriors2 %>%
      group_by(age_minbin, COD) %>%
      summarise(lcl_C_x_proportion = quantile(C_xi_proportion, 0.025),
                median_C_x_proportion = quantile(C_xi_proportion, 0.5),
                ucl_C_x_proportion = quantile(C_xi_proportion, 0.975))

    rm(age.cod.decomp.posteriors2)
  }
)
```

```{r save_results}

#SET 1: Alabama, Arizona, Arkansas, California
save(r.All, le.calculations, file = "~/BW_results/results_set_1.Rdata") 
write.csv(x = BlackWhite, file = "~/BW_results/BlackWhite_set_1.csv")
write.csv(x = age.decomp.estimates, file = "~/BW_results/age_decomp_set_1.csv")
write.csv(x = cod.marginal.estimates, file = "~/BW_results/cod_marginal_set_1.csv")
write.csv(x = age.cod.estimates, file = "~/BW_results/age_cod_set_1.csv")
```