### 2017-02-07 Code/calculate_std_mort_rates.Rmd

---
title: "Untitled"
output: html_document
---

```{r load_libraries}
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(reshape2)
library(grid)
library(gridExtra)
```

```{r}
load("~/BW_results/results_set_1.Rdata")
r.All.set1 <- r.All %>% filter(post.samp %in% c(1,2))

for(i in 2:8){
  print(i)
  load(paste0("~/BW_results/results_set_", i, ".Rdata"))
  assign(paste0("r.All.set", i), r.All %>% filter(post.samp %in% c(1,2)))
  rm(r.All)
}
```

```{r load_std_pop}
std.pop <- read.csv("~/BlackWhiteMortalityGap/Data/US_Standard_Population.csv")

std.pop <- std.pop %>% select(Age, US_Standard_2000) %>% filter(Age != "Total")

std.pop <- std.pop %>% mutate(age.weight = US_Standard_2000/1000000)
```