### 2017-04-28 Code/4_combined-meta-regression.Rmd

---
title: "meta-analysis"
author: "Corinne Riddell"
date: "4/7/2017"
output: html_document
---

In this file, we perform the meta-regression. This regression brings together the mean estimates and their 95% CIs from the state-specific models to create a smoothed line of the general pattern (across the US) of the contribution of each cause of death to the black-white life expectancy gap, separately for males and females.

In the paper, we perform a random-effects meta regression, that includes random intercepts for each state and plot the general pattern (when the RE estimate is equal to 0). In this file we also include code for the fixed effect estimation, for interested readers. This model directly estimates vertical shifts to the mean curve for each state, which may be of interest to some readers.

```{r initialize, warning=F, message=F}
library(lme4)
library(splines)
library(broom)
library(tidyverse)

source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")
```

```{r load-data}
cod_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/cod_decomp_results.csv")
cod_decomp_results$COD_cont_yrs_width <- cod_decomp_results$COD_cont_yrs_ucl - cod_decomp_results$COD_cont_yrs_lcl
cod_decomp_results$COD_cont_yrs_SE <- (cod_decomp_results$COD_cont_yrs_ucl - cod_decomp_results$COD_cont_yrs_lcl)/3.92

age_cod_results_female <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/age_cod_results_female.csv")
age_cod_results_male <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/age_cod_results_male.csv")
age_cod_results <- rbind(age_cod_results_female, age_cod_results_male)
rm(age_cod_results_female, age_cod_results_male)

age_cod_results$age_COD_cont_yrs_SE <- (age_cod_results$age_COD_cont_yrs_ucl - age_cod_results$age_COD_cont_yrs_lcl)/3.92
age_cod_results_infants_AOC <- subset(age_cod_results, age_minbin == 0 & COD == "All other causes")
```

```{r RE_meta_regression_of_cod_contribution, echo = F}
#Random effects meta analysis (fixed effects for state)

#run 12 stratified models. bs() created the b-spline basis
RE.models <- cod_decomp_results %>% 
  group_by(COD, sex) %>%
  do(fit = lmer(COD_cont_yrs_mean ~ bs(year, degree = 1, df = 5) + (1|state),
              weights = 1/(COD_cont_yrs_SE)^2, 
              data = .))

#calculated predicted values for all models
RE.pred <- broom::augment(RE.models, fit)
RE.pred$year <- 1969:2013

RE.models$RE.estimates <- purrr::map(RE.models$fit, ~ ranef(.x))

sex <- rep(c(rep("Female", 40), rep("Male", 40)), 6)
COD <- c(rep("All other causes", 80), rep("Cancers", 80), rep("Cardiovascular", 80),
         rep("Communicable", 80), rep("Injuries", 80), rep("Non-communicable", 80))
RE.estimates <- c(RE.models$RE.estimates[[1]]$state$`(Intercept)`, RE.models$RE.estimates[[2]]$state$`(Intercept)`, 
                  RE.models$RE.estimates[[3]]$state$`(Intercept)`, RE.models$RE.estimates[[4]]$state$`(Intercept)`, 
                  RE.models$RE.estimates[[5]]$state$`(Intercept)`, RE.models$RE.estimates[[6]]$state$`(Intercept)`, 
                  RE.models$RE.estimates[[7]]$state$`(Intercept)`, RE.models$RE.estimates[[8]]$state$`(Intercept)`, 
                  RE.models$RE.estimates[[9]]$state$`(Intercept)`, RE.models$RE.estimates[[10]]$state$`(Intercept)`,
                  RE.models$RE.estimates[[11]]$state$`(Intercept)`, RE.models$RE.estimates[[12]]$state$`(Intercept)`)
state <- rep(levels(cod_decomp_results$state), 12)
REs <- data.frame(sex, COD, RE.estimates, state)

cod_decomp_results <- merge(cod_decomp_results,
                            REs, by = c("COD", "sex", "state"),
                            all.x = T)

#merge the predictions with original dataset
cod_decomp_results <- merge(cod_decomp_results, 
                             RE.pred %>% select(COD, sex, state, year, .fitted, .resid, .fixed), 
                             by = c("COD", "sex", "state", "year"))

cod_decomp_results <- cod_decomp_results %>% rename(fitted.RE = .fitted, fixed.from.RE = .fixed, resid.RE = .resid)
```

```{r FE_meta_regression_of_cod_contribution, echo = F}
#Fixed effects meta analysis (fixed effects for state)

#run 12 stratified models. bs() created the b-spline basis
models <- cod_decomp_results %>% 
  group_by(COD, sex) %>%
  do(fit = lm(COD_cont_yrs_mean ~ bs(year, degree = 1, df = 5) + state,
              weights = 1/(COD_cont_yrs_SE)^2, 
              data = .))

#calculated predicted values for all models
dfPred <- broom::augment(models, fit)
dfPred$year <- 1969:2013

#merge the predictions with original dataset
cod_decomp_results <- merge(cod_decomp_results, 
                            dfPred %>% select(COD, sex, year, state, .fitted, .resid), 
                            by = c("COD", "sex", "state", "year")) %>%
                      rename(fitted.FE = .fitted, resid.FE = .resid)

#calculate the sum of the residuals for each state-sex-cod strata (summed over years)
cod_decomp_results <- cod_decomp_results %>% 
  group_by(state, sex, COD) %>%
  mutate(sum.residual = sum(abs(resid.FE)))
```

```{r RE_meta_regression_agecod_infants_only, echo = F}
RE.models.infants <- age_cod_results_infants_AOC %>% 
  group_by(sex) %>%
  do(fit = lmer(age_COD_cont_yrs_mean ~ bs(year, degree = 1, df = 5) + (1|state),
              weights = 1/(age_COD_cont_yrs_SE)^2, 
              data = .))

#creates a dataframe for males and females of the fitted estimates. Includes state variable, but we need to add on the year variable
RE.pred.infants <- broom::augment(RE.models.infants, fit)
RE.pred.infants$year <- 1969:2013

RE.models.infants$RE.estimates <- purrr::map(RE.models.infants$fit, ~ ranef(.x))

sex.infants <- c(rep("Female", 40), rep("Male", 40))
COD.infants <- rep("All other causes", 80)
RE.estimates.infants <- c(RE.models.infants$RE.estimates[[1]]$state$`(Intercept)`, RE.models.infants$RE.estimates[[2]]$state$`(Intercept)`)
state.infants <- rep(levels(age_cod_results$state), 2)
REs.infants <- data.frame(sex.infants, COD.infants, RE.estimates.infants, state.infants)
REs.infants <- REs.infants %>% rename(sex = sex.infants, COD = COD.infants, RE.estimates = RE.estimates.infants, state = state.infants)

#merge in the RE estimates to the original dataset
age_cod_results_infants_AOC <- merge(age_cod_results_infants_AOC,
                            REs.infants, by = c("sex", "state"), #state-specific REs; don't need time for merge (because time fixed)
                            all.x = T)

#merge other aspects, like the FE, with original dataset
age_cod_results_infants_AOC <- merge(age_cod_results_infants_AOC, 
                             RE.pred.infants %>% select(sex, state, year, .fitted, .resid, .fixed), 
                             by = c("sex", "state", "year"))

age_cod_results_infants_AOC <- age_cod_results_infants_AOC %>% rename(fitted.RE = .fitted, fixed.from.RE = .fixed, resid.RE = .resid)
```

```{r save_data}
write.csv(cod_decomp_results, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/cod_decomp_results.csv")

write.csv(age_cod_results_infants_AOC, "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/age_cod_results_infants_AOC.csv")
```