### 2017-05-03 conferences/ResearchDay/Riddell_ResearchDay_minimal.Rmd
research day stuff

---
title: "Changes in cause of death and the black-white life expectancy gap"
author: "Corinne Riddell, PhD"
date: "April 24, 2017"
output: ioslides_presentation
---

<head>
<style>
.greentext {
        color: green;
}
.redtext {
        color: red;
}
</style>
</head>

```{r initialize, echo=FALSE, warning=F, message=F}
library(plotly)
library(viridis)
library(scales)
library(dplyr)
library(purrr)
library(stringr)
library(DT)
library(splines)
library(broom)
library(tidyverse)

library(png)
library(jpeg)
library(grid)
library(gridExtra)
library(knitr)

source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")
```

## Thank you

```{r, echo=F}
img1 <-  rasterGrob(as.raster(readPNG("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/ResearchDay/images/k_morrison.png")), interpolate = FALSE)
img2 <-  rasterGrob(as.raster(readPNG("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/ResearchDay/images/s_harper2.png")), interpolate = FALSE)
img3 <-  rasterGrob(as.raster(readJPEG("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/ResearchDay/images/j_kaufman.jpg")), interpolate = FALSE)
grid.arrange(img3,
             img1,
             img2,
             ncol = 3)
```

----

```{r, echo = F, out.width = 500}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/ResearchDay/images/Harper_HA_title.png")
```

```{r, echo=F, out.width = 750}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/ResearchDay/images/Harper_HA_fig1.png")
```

----

```{r, echo = F, out.width = 250}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/ResearchDay/images/Harper_HA_quote.png")
```

And also...

>- be broadly accessible,
>- be interactive,
>- be reproducible,
>- and allow pictures to speak for themselves.

## The data

```{r, echo = F, out.width = 400}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/ResearchDay/images/alabama.data.ex.png")
```

- <b class = "greentext">40</b> "states": 39 states + DC
- <b class = "greentext">45</b> years: 1969--2013 
- <b class = "greentext">19</b> age groups: <1, 1-4, 5-9, ..., 85+
- <b class = "greentext">2</b> sexes
- <b class = "greentext">2</b> races (Black, White)
- <b class = "greentext">6</b> major causes of deaths: CVD, cancer, non-communicable, communicable, injuries, all other causes
- = <b class = "greentext">820,800 rows of data</b>

## The data

```{r, echo = F, out.width = 400}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/ResearchDay/images/alabama.data.ex.png")
```

- When the number of deaths is between 1 and 9 it is <b class = "redtext">coarsened</b>.

## The analysis

**Estimate smoothed mortality rates and impute coarsened data.**

+ Bayesian autoregressive model to borrow information *over time* within strata
+ $log(\lambda_t) \sim N(\ log(\lambda_{t-1}),\ \tau)$
+ impute coarsened 1--9 values using a truncated Poisson
+ `rjags`

## The analysis

**Create 1,000 posterior samples**

+ Each sample contains estimated *smoothed* and *imputed* death counts
+ For each sample:
    1. Create a life table for every strata
    2. Calculate the <b style = "greentext">difference</b> in life expectancy between blacks and whites
    3. Decompose the difference by age and cause of death
    4. Calculate age-standardized mortality rates
+ Calculate the mean, 5^th^ and 95^th^ percentiles for all estimates.

## The analysis

**Synthesize state-specific findings**

+ Random-effects meta regression to estimate overall trend in contributions for each cause.
+ Use `ggplot2`, `plotly`, and `Shiny` to make results accessible as a web app.

----
# To the results!