### 2017-06-02 conferences/Riddell_presentation_max_planck.Rmd
moved conference files

---
title: "Trends in the contribution of major causes of death to the black-white life expectancy gap"
author: "Corinne Riddell, Post-doctoral Researcher, McGill University, Canada"
date: "May 23, 2017"
output: ioslides_presentation
---

<head>
<style>
.greentext {
        color: green;
}
.redtext {
        color: red;
}
</style>
</head>

```{r initialize, echo=FALSE, warning=F, message=F}
library(plotly)
library(viridis)
library(scales)
library(dplyr)
library(purrr)
library(stringr)
library(DT)
library(splines)
library(broom)
library(tidyverse)

library(png)
library(jpeg)
library(grid)
library(gridExtra)
library(knitr)

source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")
BlackWhite_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/BlackWhite_results.csv")
BlackWhite_results <- reorder.as.map2(BlackWhite_results, "state", "stabbrs")

```

## Thank you

```{r, echo=F}
img1 <-  rasterGrob(as.raster(readPNG("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/k_morrison.png")), interpolate = FALSE)
img2 <-  rasterGrob(as.raster(readPNG("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/s_harper2.png")), interpolate = FALSE)
img3 <-  rasterGrob(as.raster(readJPEG("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/j_kaufman.jpg")), interpolate = FALSE)
grid.arrange(img3,
             img1,
             img2,
             ncol = 3)
```

Jay Kaufman, Kathryn Morrison, and Sam Harper

----

```{r, echo = F, out.width = 500}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/Harper_HA_title.png")
```

```{r, echo=F, out.width = 750}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/Harper_HA_fig1.png")
```

----

```{r, echo = F, out.width = 250}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/Harper_HA_quote.png")
```

## The data

```{r, echo = F, out.width = 400}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/alabama.data.ex.png")
```

- <b class = "greentext">40</b> "states": 39 states + DC
- <b class = "greentext">45</b> years: 1969--2013 
- <b class = "greentext">19</b> age groups: <1, 1-4, 5-9, ..., 85+
- <b class = "greentext">2</b> genders
- <b class = "greentext">2</b> races (black, white)
- <b class = "greentext">6</b> major causes of deaths: cardiovascular disease, cancer, non-communicable, communicable, injuries, all other causes
- = <b class = "greentext">820,800 rows of data</b>

## The analysis

+ Estimated smoothed mortality rates using a Bayesian time series model with Poisson likelihood
+ Used a truncated Poisson likelihood when number of deaths were between 1 and 9
+ 1,000 posterior samples. For each:
    1. Create a life table for every strata
    2. Calculate the <b style = "greentext">difference</b> in life expectancy between blacks and whites
    3. Decompose the difference by age and cause of death
    4. Calculate age-standardized mortality rates
+ Calculate the mean, 5^th^ and 95^th^ percentiles for all estimates
+ Compute national trend using random-effects meta regression

## The output

+ Working paper being screened at bioRxiv
+ R Shiny App: https://corinne-riddell.shinyapps.io/shiny_app/

## The app

```{r, echo = F, out.width = 800}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/app_ex1.gif")
```

## The app

```{r, echo = F, out.width = 800}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/app_ex2.gif")
```

## The app

```{r, echo = F, out.width = 800}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/app_ex3.gif")
```

## Male black-white life expectancy gap

```{r, echo = F}
    gg.p <- ggplot(data = subset(BlackWhite_results, sex == "Male"),
                          aes(y = LE_wbgap_mean, x = year)) +
    geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) +
    geom_ribbon(aes(ymin = LE_wbgap_lcl, ymax = LE_wbgap_ucl, fill = Census_Division)) +
    #geom_line(aes(col = Census_Division)) + 
    facet_wrap(~ stabbrs.map.order, ncol = 11, drop = F) +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank(),
          strip.background=element_blank(),
          axis.line=element_blank(),
          axis.ticks=element_blank(),
          legend.position = "none",
          panel.background = element_rect(fill = "transparent", colour = NA), 
          plot.background = element_rect(fill = "transparent", colour = NA),
          legend.background = element_rect(fill = "transparent", colour = NA)) +
    ylab("Life expectancy gap (years)") +
    xlab(paste0("Year (1969-2013)"))

      interactive.p <- ggplotly(gg.p +     
        geom_hline(aes(yintercept = 5), lwd = 0.5, col = "grey", alpha = 0.5) +
        geom_hline(aes(yintercept = 10), lwd = 0.5, col = "grey", alpha = 0.5) +
        geom_hline(aes(yintercept = 15), lwd = 0.5, col = "grey", alpha = 0.5))

      for(i in 1:length(interactive.p$x$data)){
        if (interactive.p$x$data[[i]]$line$color %in% c("rgba(0,0,0,1)", "rgba(190,190,190,0.5)")) {
          interactive.p$x$data[[i]]$hoverinfo <- "none"
          interactive.p$x$data[[i]]$line$width <- 1
        }

        interactive.p$x$data[[i]]$text <- gsub("Census_Division", "Census Division", interactive.p$x$data[[i]]$text)
        interactive.p$x$data[[i]]$text <- gsub("LE_wbgap_mean", "Mean life expectancy gap", interactive.p$x$data[[i]]$text)
        interactive.p$x$data[[i]]$text <- gsub("LE_wbgap_ucl", "Upper credible limit", interactive.p$x$data[[i]]$text)
        interactive.p$x$data[[i]]$text <- gsub("LE_wbgap_lcl", "Lower credible limit", interactive.p$x$data[[i]]$text)
      }

       interactive.p
```

## Female black-white life expectancy gap

```{r, echo = F}
    gg.p2 <- ggplot(data = subset(BlackWhite_results, sex == "Female"),
                          aes(y = LE_wbgap_mean, x = year)) +
    geom_hline(aes(yintercept = 0)) + geom_vline(aes(xintercept = 1969)) +
    geom_ribbon(aes(ymin = LE_wbgap_lcl, ymax = LE_wbgap_ucl, fill = Census_Division)) +
    #geom_line(aes(col = Census_Division)) + 
    facet_wrap(~ stabbrs.map.order, ncol = 11, drop = F) +
    theme_classic(base_size = 10) +
    theme(axis.text.x = element_blank(),
          strip.background=element_blank(),
          axis.line=element_blank(),
          axis.ticks=element_blank(),
          legend.position = "none",
          panel.background = element_rect(fill = "transparent", colour = NA), 
          plot.background = element_rect(fill = "transparent", colour = NA),
          legend.background = element_rect(fill = "transparent", colour = NA)) +
    ylab("Life expectancy gap (years)") +
    xlab(paste0("Year (1969-2013)"))

      interactive.p2 <- ggplotly(gg.p2 +     
        geom_hline(aes(yintercept = -5), lwd = 0.5, col = "grey", alpha = 0.5) +
        geom_hline(aes(yintercept = 5), lwd = 0.5, col = "grey", alpha = 0.5) +
        geom_hline(aes(yintercept = 10), lwd = 0.5, col = "grey", alpha = 0.5))

      for(i in 1:length(interactive.p2$x$data)){
        if (interactive.p2$x$data[[i]]$line$color %in% c("rgba(0,0,0,1)", "rgba(190,190,190,0.5)")) {
          interactive.p2$x$data[[i]]$hoverinfo <- "none"
          interactive.p2$x$data[[i]]$line$width <- 1
        }

        interactive.p2$x$data[[i]]$text <- gsub("Census_Division", "Census Division", interactive.p$x$data[[i]]$text)
        interactive.p2$x$data[[i]]$text <- gsub("LE_wbgap_mean", "Mean life expectancy gap", interactive.p$x$data[[i]]$text)
        interactive.p2$x$data[[i]]$text <- gsub("LE_wbgap_ucl", "Upper credible limit", interactive.p$x$data[[i]]$text)
        interactive.p2$x$data[[i]]$text <- gsub("LE_wbgap_lcl", "Lower credible limit", interactive.p$x$data[[i]]$text)
      }

       interactive.p2
```

##National patterns in cause contributions to the gap

```{r, echo = F}
cod_decomp_results <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/cod_decomp_results.csv")
cod_decomp_results$COD <- factor(cod_decomp_results$COD, levels(cod_decomp_results$COD)[c(3, 2, 4, 6, 5, 1)])

cod_decomp_national <- cod_decomp_results %>% 
  filter(state == "Alabama") %>%
  mutate(national.trend = fitted.RE - RE.estimates) %>%
  select(-state)

gg.p3 <- ggplotly(ggplot(cod_decomp_national, aes(y = national.trend, x = year)) + 
  geom_line(aes(col = sex)) + 
  facet_wrap(~COD) + theme_minimal() +
  xlab("Year") + ylab("National trend in cause of death contribution")) %>%
  layout(margin(b = 50, l = 50))

gg.p3
```

#Three examples

##CVD in men
```{r, echo = F, out.width = 675}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/CVD_men.png")
```

##Cancer in women
```{r, echo = F, out.width = 675}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/Cancer_women.png")
```

##Injury in women
```{r, echo = F, out.width = 675}
include_graphics(path = "/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/conferences/images/Injury_women.png")
```

#A tale of two states

##Tennessee
```{r, echo = F, out.height = 600, out.width = 800}
age_cod_results_female <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/age_cod_results_female.csv")

age_cod_results_male <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results2/age_cod_results_male.csv")

age_cod_results <- rbind(age_cod_results_female, age_cod_results_male)
rm(age_cod_results_female, age_cod_results_male)

age_cod_results$age <- forcats::fct_relevel(f = age_cod_results$age, "<1 year", "1-4 years", "5-9 years", "10-14 years", 
                                            "15-19 years", "20-24 years", "25-29 years", "30-34 years", "35-39 years", 
                                            "40-44 years", "45-49 years", "50-54 years", "55-59 years", "60-64 years", 
                                            "65-69 years", "70-74 years", "75-79 years", "80-84 years", "85+ years")

age_cod_results$COD <- factor(age_cod_results$COD, levels(age_cod_results$COD)[c(3, 2, 4, 6, 5, 1)])

sub1 <- subset(age_cod_results, year %in% c(1969, 2013) & state == "Tennessee")
ly1 <- ggplotly(
  ggplot(data = sub1,
         aes(x=age, y = age_COD_cont_yrs_mean, fill = COD)) +
    geom_bar(stat = "identity") + coord_flip() + theme_minimal()  + geom_vline(aes(xintercept = 0)) +
    theme(legend.title = element_blank(),
          panel.background = element_rect(fill = "transparent", colour = NA), 
          plot.background = element_rect(fill = "transparent", colour = NA),
          legend.background = element_rect(fill = "transparent", colour = NA)) +
    ylab("Contribution to the life expectancy gap (years)") +
    xlab("") +
    facet_wrap(year ~ sex) 
)

for(i in 1:length(ly1$x$data)){
  ly1$x$data[[i]]$text <- gsub("age_COD_cont_yrs_mean", "Contribution to gap (yrs)", ly1$x$data[[i]]$text)
}

ly1 %>% layout(margin = list(b = 100))
```

##New York
```{r, echo = F, out.height = 600, out.width = 800}

sub1 <- subset(age_cod_results, year %in% c(1969, 2013) & state == "New York")
ly1 <- ggplotly(
  ggplot(data = sub1,
         aes(x=age, y = age_COD_cont_yrs_mean, fill = COD)) +
    geom_bar(stat = "identity") + coord_flip() + theme_minimal()  + geom_vline(aes(xintercept = 0)) +
    theme(legend.title = element_blank(),
          panel.background = element_rect(fill = "transparent", colour = NA), 
          plot.background = element_rect(fill = "transparent", colour = NA),
          legend.background = element_rect(fill = "transparent", colour = NA)) +
    ylab("Contribution to the life expectancy gap (years)") +
    xlab("") +
    facet_wrap(year ~ sex) 
)

for(i in 1:length(ly1$x$data)){
  ly1$x$data[[i]]$text <- gsub("age_COD_cont_yrs_mean", "Contribution to gap (yrs)", ly1$x$data[[i]]$text)
}

ly1 %>% layout(margin = list(b = 100))
```

##Similarities & differences

- Substantial contribution of infant mortality to the gap in 1969
- Most deaths from all other causes are during infancy
- Similar pattern of injury contributions by gender in 1969, but larger contribution in Tennessee
- Similar pattern of CVD contributions in 1969 and 2013, but Tennessee started worse off
- Ditto for cancer
- Injury plays important role in Tennessee in 2013 with opposing effect directions in men and women

##Get in touch
- corinne.riddell@mail.mcgill.ca
- @datavisitor on Twitter