### 2017-06-02 Code/archived/Run_model_on_SH_data/Import_SH_HA_data.Rmd
reorg

---
title: "extract table from SH HA paper"
author: "Corinne Riddell"
date: "January 22, 2017"
output: html_document
---

```{r}

library(pdftools)

txt <- pdf_text("/Users/corinneriddell/Dropbox/BlackWhiteGap/papers/Harper-Health Affairs-2014.pdf")

#males
cat(txt[13])
cat(txt[14])

#females
cat(txt[15])
cat(txt[16])

#i then copied and pasted the above into excel and saved as SH_HA_supp_app_data.csv
#to bring back here...

SH.dat <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Data/SH_HA_supp_app_data.csv", header = T)
```