### 2017-06-06 Code/5_RI_sensitivity.Rmd
CR sensitivity analysis

---
title: "Rhode_Island_sensitivity"
output: html_document
---

```{r setup, include=FALSE}
dat.clean = read.csv('~/dat_clean.csv')

RI <- dat.clean %>% filter(state == "Rhode Island")

RI.1999 <- RI %>% filter(year >= 1999)

rm(dat.clean)
```

```{r}
entire.analysis <- function(state_i, chosen.seed) {

  #load libraries
  library(Rcpp)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(stringr)
  library(rjags)
  library(R2jags) 
  library(devtools)
  library(jagstools)
  library(reshape2)
  library(parallel)
  library(gtools)
  library(bayesboot)
  library(coda)
  library(grid)
  library(gridExtra)
  library(parallel)

  #load user written functions
  source("~/repos/BlackWhiteMortalityGap/Code/2_smoothing_models/bayes_smoothing_functions.R")
  source('~/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R')

  #load tidied data
  dat.clean = RI.1999

  #load US standard population
  std.pop <- read.csv("~/repos/BlackWhiteMortalityGap/Data/US_Standard_Population_for_model.csv") %>% select(US_Standard_2000, age.weight, age_minbin)

  #fit models
  #run time: 32 mins
  #fit the bayesian models and put the 1000 posterior samples into a data frame
  for(sex_i in c("Male", "Female")){
    print(sex_i)
    assign(paste0("r.White.", sex_i), run_analysis(dataset = dat.clean, state1 = state_i, sex1 = sex_i, race1 = "White", chosen.seed))
    assign(paste0("r.Black.", sex_i), run_analysis(dataset = dat.clean, state1 = state_i, sex1 = sex_i, race1 = "Black", chosen.seed))
  }

  r.White.Male$race <- "White"
  r.White.Female$race <- "White"
  r.Black.Male$race <- "Black"
  r.Black.Female$race <- "Black"
  r.All <- rbind(r.White.Male, r.White.Female, r.Black.Male, r.Black.Female)
  rm(r.White.Male, r.White.Female, r.Black.Male, r.Black.Female)

  #calculate age-standardized mortality rates (for each posterior sample)
  r.All <- merge(r.All, std.pop, by = "age_minbin")

  mortality.rates <- r.All %>% 
    group_by(post.samp, state.n, year.n, race.n, sex.n, COD) %>% 
    arrange(age_minbin) %>%
    summarise(age.std.mortality.rate = sum(age.weight*smoothed_rate), #the sum is across age bins
              rate.per.100k = age.std.mortality.rate*100000,
              race = first(race),
              sex = first(sex),
              state = first(state),
              year = first(year))

  mortality.rates <- add_Census_Division(mortality.rates, state)
  mortality.rates <- add_Census_Region(mortality.rates, state)
  mortality.rates$stabbrs <- state.abb[match(mortality.rates$state, state.name)]

  #calculate median, mean, and CI for age-standardized mortality rates (across samples) 
  mortality.rates.summary <- mortality.rates %>% group_by(state.n, year.n, race.n, sex.n, COD) %>%
    summarise(race = first(race),
              sex = first(sex),
              state = first(state),
              year = first(year),
              Census_Division = first(Census_Division),
              Census_Region = first(Census_Region),
              stabbrs = first(stabbrs),
              rate.per.100k_lcl = quantile(rate.per.100k, 0.025), 
              rate.per.100k_med = quantile(rate.per.100k, 0.5),
              rate.per.100k_ucl = quantile(rate.per.100k, 0.975),
              rate.per.100k_mean = mean(rate.per.100k))

  #compute black-white difference in mortality
  mortality.rates.white <- mortality.rates %>% filter(race == "White") %>% 
    ungroup() %>%
    select(sex, state, year, COD, post.samp, rate.per.100k) %>%
    rename(rate.per.100k.white = rate.per.100k)

  mortality.rates.black <- mortality.rates %>% filter(race == "Black") %>% 
    rename(rate.per.100k.black = rate.per.100k)

  mortality.rates.wide <- merge(mortality.rates.black, mortality.rates.white, by = c("post.samp", "sex", "state", "year", "COD"))

  mortality.rates.wide <- mortality.rates.wide %>% mutate(rate.difference = rate.per.100k.black- rate.per.100k.white)

  mortality.rates.difference <- mortality.rates.wide %>% 
    group_by(state.n, year.n, sex.n, COD) %>%
    summarise(rate.difference_LCL = quantile(rate.difference, 0.025), #summarize across posterior samples (post.samp)
              rate.difference_med = quantile(rate.difference, 0.5),
              rate.difference_UCL = quantile(rate.difference, 0.975),
              rate.difference_mean = mean(rate.difference),
              Census_Division = first(Census_Division),
              Census_Region = first(Census_Region),
              stabbrs = first(stabbrs),
              sex = first(sex),
              state = first(state),
              year = first(year))

  r.All$smoothed_deaths <- r.All$smoothed_rate*r.All$population
  table(r.All$smoothed_deaths == 0) #check this -- if no deaths in an age-band (i.e., no death for any COD, then can't run the COD calculations for that stratum.)

  r.All <- r.All %>% group_by(post.samp, state, race.sex, year.n, age_minbin) %>%
    arrange(post.samp, state, race.sex, year.n, age_minbin)

  #RUN LIFE EXPECTANCY AND DECOMPOSITION CALCULATIONS
  #run time: 2.25 hours per state
  #run the life table and decomposition calculations for each of the 1000 posterior samples

  strata <- levels(droplevels(r.All$state.year.sex))
  n <- length(strata)

  le.calculations = vector(mode = "list", length = n)

  counter = 1

  for(stratum_i in strata){
    data_sub <- subset(r.All, state.year.sex == stratum_i)

    le.calculations[[counter]]$stratum.id <- stratum_i
    le.calculations[[counter]]$calcs <- by(data = data_sub,
                                           INDICES = list(data_sub$post.samp),
                                           FUN = function(x) life_expectancy_and_gap(data = x))

    counter <- counter + 1

    print(stratum_i)
    rm(data_sub)
  }

  save(r.All, file = paste0("~/RhodeIsland/r_All_", state_i, "_", chosen.seed, "_1999.Rdata")) 
  levels.state.year.sex <- levels(droplevels(r.All$state.year.sex))
  rm(r.All)

  #extract_life_expectancy_results
  BlackWhite <- data.frame(stratum.id = character(),
                           LE_white_lcl = numeric(), LE_white_med = numeric(), LE_white_ucl = numeric(), LE_white_mean = numeric(), 
                           LE_black_lcl = numeric(), LE_black_med = numeric(), LE_black_ucl = numeric(), LE_black_mean = numeric(), 
                           LE_wbgap_lcl = numeric(), LE_wbgap_med = numeric(), LE_wbgap_ucl = numeric(), LE_wbgap_mean = numeric(), 
                           stringsAsFactors = F)

  age_minbin <- le.calculations[[1]]$calcs[[1]]$age.decomp$Ages
  cods <- le.calculations[[1]]$calcs[[1]]$cod.marginal$Cause.of.death

  age.decomp.estimates <- data.frame(stratum.id = rep(levels.state.year.sex, 
                                                      each = length(age_minbin)),
                                     age_minbin = numeric(19*n),
                                     age_cont_yrs_lcl = numeric(19*n), 
                                     age_cont_yrs_med = numeric(19*n), 
                                     age_cont_yrs_ucl = numeric(19*n),
                                     age_cont_yrs_mean = numeric(19*n), 
                                     age_cont_prop_lcl = numeric(19*n), 
                                     age_cont_prop_med = numeric(19*n), 
                                     age_cont_prop_ucl = numeric(19*n),
                                     age_cont_prop_mean = numeric(19*n), 
                                     stringsAsFactors = F)

  cod.marginal.estimates <- data.frame(stratum.id = rep(levels.state.year.sex, 
                                                        each = length(cods)),
                                       COD = character(6*n),
                                       COD_cont_yrs_lcl = numeric(6*n), 
                                       COD_cont_yrs_med = numeric(6*n), 
                                       COD_cont_yrs_ucl = numeric(6*n),
                                       COD_cont_yrs_mean = numeric(6*n), 
                                       COD_cont_prop_lcl = numeric(6*n), 
                                       COD_cont_prop_med = numeric(6*n), 
                                       COD_cont_prop_ucl = numeric(6*n),
                                       COD_cont_prop_mean = numeric(6*n), 
                                       stringsAsFactors = F)

  age.cod.estimates <- data.frame(stratum.id = rep(levels.state.year.sex, 
                                                   each = length(cods)*length(age_minbin)),
                                  age_minbin = numeric(19*6*n),   
                                  COD = character(19*6*n),
                                  age_COD_cont_yrs_lcl = numeric(19*6*n), 
                                  age_COD_cont_yrs_med = numeric(19*6*n), 
                                  age_COD_cont_yrs_ucl = numeric(19*6*n),
                                  age_COD_cont_yrs_mean = numeric(19*6*n),
                                  age_COD_cont_prop_lcl = numeric(19*6*n), 
                                  age_COD_cont_prop_med = numeric(19*6*n), 
                                  age_COD_cont_prop_ucl = numeric(19*6*n), 
                                  age_COD_cont_prop_mean = numeric(19*6*n), 
                                  stringsAsFactors = F)

  for(i in 1:n) {

    BlackWhite[i, c("stratum.id")] <- le.calculations[[i]]$stratum.id

    #LIFE EXPECTANCY EXTRACT RESULTS
    BlackWhite[i , c("LE_white_lcl", "LE_white_med", "LE_white_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                          ~.x$le.df$LE_White), 
                                                                                  c(0.025, 0.5, 0.975))
    BlackWhite[i , c("LE_black_lcl", "LE_black_med", "LE_black_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                          ~.x$le.df$LE_Black), 
                                                                                  c(0.025, 0.5, 0.975))
    BlackWhite[i , c("LE_wbgap_lcl", "LE_wbgap_med", "LE_wbgap_ucl")] <- quantile(map_dbl(le.calculations[[i]]$calcs,
                                                                                          ~.x$le.df$LE_WBgap), 
                                                                                  c(0.025, 0.5, 0.975))

    BlackWhite[i , c("LE_white_mean")] <- mean(map_dbl(le.calculations[[i]]$calcs, ~.x$le.df$LE_White))
    BlackWhite[i , c("LE_black_mean")] <- mean(map_dbl(le.calculations[[i]]$calcs, ~.x$le.df$LE_Black))
    BlackWhite[i , c("LE_wbgap_mean")] <- mean(map_dbl(le.calculations[[i]]$calcs, ~.x$le.df$LE_WBgap))

    #AGE DECOMPOSITION EXTRACT RESULTS
    age.decomp.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$age.decomp$C_x)
    age.decomp.posteriors$age_minbin <- age_minbin
    age.decomp.posteriors <- age.decomp.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x")

    age.decomp.estimates[age.decomp.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                         c("age_minbin", "age_cont_yrs_lcl", "age_cont_yrs_med", "age_cont_yrs_ucl", "age_cont_yrs_mean")] <- 
      age.decomp.posteriors %>% 
      group_by(age_minbin) %>%
      summarise(lcl_C_x = quantile(C_x, 0.025),
                median_C_x = quantile(C_x, 0.5), 
                ucl_C_x = quantile(C_x, 0.975),
                mean_C_x = mean(C_x))

    rm(age.decomp.posteriors)

    age.decomp.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$age.decomp$C_x_proportion)
    age.decomp.posteriors2$age_minbin <- age_minbin
    age.decomp.posteriors2 <- age.decomp.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x_proportion")

    age.decomp.estimates[age.decomp.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                         c("age_minbin", "age_cont_prop_lcl", "age_cont_prop_med", "age_cont_prop_ucl", "age_cont_prop_mean")] <- 
      age.decomp.posteriors2 %>% 
      group_by(age_minbin) %>%
      summarise(lcl_C_x_proportion = quantile(C_x_proportion, 0.025),
                median_C_x_proportion = quantile(C_x_proportion, 0.5), 
                ucl_C_x_proportion = quantile(C_x_proportion, 0.975),
                mean_C_x_proportion = mean(C_x_proportion))

    rm(age.decomp.posteriors2)

    #COD DECOMPOSITION EXTRACT RESULTS
    cod.marginal.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.marginal$C_x)
    cod.marginal.posteriors$COD <- as.character(cods)
    cod.marginal.posteriors <- cod.marginal.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_x")

    cod.marginal.estimates[cod.marginal.estimates$stratum.id == le.calculations[[i]]$stratum.id, 
                           c("COD", "COD_cont_yrs_lcl", "COD_cont_yrs_med", "COD_cont_yrs_ucl", "COD_cont_yrs_mean")] <- 
      cod.marginal.posteriors %>% 
      group_by(COD) %>%
      summarise(lcl_C_x = quantile(C_x, 0.025),
                median_C_x = quantile(C_x, 0.5), 
                ucl_C_x = quantile(C_x, 0.975),
                mean_C_x = mean(C_x))

    rm(cod.marginal.posteriors)

    cod.marginal.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.marginal$C_x_proportion)
    cod.marginal.posteriors2$COD <- as.character(cods)
    cod.marginal.posteriors2 <- cod.marginal.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", 
                                                                           value = "C_x_proportion")

    cod.marginal.estimates[cod.marginal.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                           c("COD", "COD_cont_prop_lcl", "COD_cont_prop_med", "COD_cont_prop_ucl", "COD_cont_prop_mean")] <-
      cod.marginal.posteriors2 %>%
      group_by(COD) %>%
      summarise(lcl_C_x_proportion = quantile(C_x_proportion, 0.025),
                median_C_x_proportion = quantile(C_x_proportion, 0.5),
                ucl_C_x_proportion = quantile(C_x_proportion, 0.975),
                mean_C_x_proportion = mean(C_x_proportion))

    rm(cod.marginal.posteriors2)

    #AGE-COD EXTRACT RESULTS
    age.cod.decomp.posteriors <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.decomp$C_xi)
    age.cod.decomp.posteriors$age_minbin <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Ages
    age.cod.decomp.posteriors$COD <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Cause.of.death
    age.cod.decomp.posteriors <- age.cod.decomp.posteriors %>% tidyr::gather(`1`:`1000`, key = "post.samp", value = "C_xi")

    age.cod.estimates[age.cod.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                      c("age_minbin", "COD", "age_COD_cont_yrs_lcl", "age_COD_cont_yrs_med", "age_COD_cont_yrs_ucl", "age_COD_cont_yrs_mean")] <-
      age.cod.decomp.posteriors %>%
      group_by(age_minbin, COD) %>%
      summarise(lcl_C_x = quantile(C_xi, 0.025),
                median_C_x = quantile(C_xi, 0.5),
                ucl_C_x = quantile(C_xi, 0.975),
                mean_C_x = mean(C_xi))

    rm(age.cod.decomp.posteriors)

    age.cod.decomp.posteriors2 <- map_df(le.calculations[[i]]$calcs, ~ .x$cod.decomp$C_xi_proportion)
    age.cod.decomp.posteriors2$age_minbin <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Ages
    age.cod.decomp.posteriors2$COD <- le.calculations[[i]]$calcs[[1]]$cod.decomp$Cause.of.death
    age.cod.decomp.posteriors2 <- age.cod.decomp.posteriors2 %>% tidyr::gather(`1`:`1000`, key = "post.samp", 
                                                                               value = "C_xi_proportion")

    age.cod.estimates[age.cod.estimates$stratum.id == le.calculations[[i]]$stratum.id,
                      c("age_minbin", "COD", "age_COD_cont_prop_lcl", "age_COD_cont_prop_med", "age_COD_cont_prop_ucl", "age_COD_cont_prop_mean")] <-
      age.cod.decomp.posteriors2 %>%
      group_by(age_minbin, COD) %>%
      summarise(lcl_C_x_proportion = quantile(C_xi_proportion, 0.025),
                median_C_x_proportion = quantile(C_xi_proportion, 0.5),
                ucl_C_x_proportion = quantile(C_xi_proportion, 0.975),
                mean_C_x_proportion = mean(C_xi_proportion))

    rm(age.cod.decomp.posteriors2)
  }

  #save_results
  save(mortality.rates, mortality.rates.wide, le.calculations,
       file = paste0("~/RhodeIsland/results_", state_i, "_", chosen.seed, "1999.Rdata")) 
  write.csv(x = BlackWhite, file = paste0("~/repos/BlackWhiteMortalityGap/sensitivity/BlackWhite_", state_i, "_", 1999,".csv"))
  write.csv(x = age.decomp.estimates, file = paste0("~/repos/BlackWhiteMortalityGap/sensitivity/age_decomp_", state_i, "_", 1999,".csv"))
  write.csv(x = cod.marginal.estimates, file = paste0("~/repos/BlackWhiteMortalityGap/sensitivity/cod_marginal_", state_i, "_", 1999,".csv"))
  write.csv(x = age.cod.estimates, file = paste0("~/repos/BlackWhiteMortalityGap/sensitivity/age_cod_", state_i, "_", 1999, ".csv"))
  write.csv(x = mortality.rates.summary, file = paste0("~/repos/BlackWhiteMortalityGap/sensitivity/mortality_rates_", state_i, "_", 1999, ".csv"))
  write.csv(x = mortality.rates.difference, file = paste0("~/repos/BlackWhiteMortalityGap/sensitivity/mortality_rates_diff_", state_i, "_", 1999, ".csv"))
}
```

```{r run_analysis}
system.time(entire.analysis(state_i = "Rhode Island", chosen.seed = 123))
```