### 2018-02-25 Code/6-Excluded-states-description.Rmd

---
title: "Excluded states description"
author: "Corinne Riddell"
date: "2/21/2018"
output: html_document
---

```{r}
library(tidyverse)
library(ggrepel)
```

```{r}
dat.aggregated <- readr::read_csv("./Data/dat_aggregated.csv", col_names = T)

dat2 <- dat.aggregated %>% 
  group_by(state, year) %>% 
  filter(age == "<1 year") %>% 
  select(state, year, race, sex, race.sex, pop_across_age)

dat3 <- dat2 %>% 
  group_by(state, year, race) %>% 
  summarise(pop.total.across.race = sum(pop_across_age))

dat3 <- dat2 %>% 
  group_by(state, year, race) %>% 
  summarise(pop.total.across.race = sum(pop_across_age),
            prop.race = round((pop.total.across.race/pop.total)*100, 1) )

dat4 <- dat3 %>% filter(state %in% c("Alaska", "Hawaii", "Idaho", "Maine", "Montana", "New Hampshire", "North Dakota", "South Dakota", "Utah", "Vermont", "Wyoming"))

```

```{r}

ggplot(dat3 %>% filter(race == "Black"), aes(year, prop.race)) + 
  geom_line(aes(col = state, group = state)) + 
  ylab("% of population that is black") + 
  geom_text_repel(data = dat4 %>% filter(race == "Black", year == 2013), aes(group = state, label = state), size = 3) + 
  ggtitle("Proportion of blacks in the population")

ggplot(dat3 %>% filter(race == "Black"), aes(year, pop.total.across.race)) + 
  geom_line(aes(col = state, group = state)) + 
  ylab("% of population that is black") + 
  geom_text_repel(data = dat4 %>% filter(race == "Black", year == 2013), aes(group = state, label = state), size = 3) + 
  ggtitle("Absolute size of black population") + guides(col = F) + 
  scale_y_log10(breaks = c(10000, 100000, 500000, 1000000, 3000000), 
                      labels = c("10k", "100k", "500k", "1m", "3m")) +
  # scale_y_continuous(breaks = c(10000, 100000, 500000, 1000000, 3000000), 
  #                    labels = c("10k", "100k", "500k", "1m", "3m")) +
  theme(panel.grid.minor = element_blank())

```